### 2025-04-22 appendix.qmd
Create manuscript file

\vspace*{60mm}

# **Supplementary Materials**

\setcounter{page}{1}
\begin{center}
Negative Perceptions of Outsourcing to Artificial Intelligence \\
Scott Claessens, Pierce Veitch, and Jim Everett \\
School of Psychology, University of Kent
\end{center}
\newpage

# Pilot Study 1

## Methods

### Participants

```{r}
pilot_sample <-
  tar_read(pilot_data) %>%
  group_by(id) %>%
  summarise(
    gender = unique(gender),
    age = unique(age)
    )
```

We recruited a convenience sample of 200 participants from the United Kingdom
through Prolific. After excluding participants who failed our pre-treatment 
attention check, we were left with a final sample of 
`r nrow(pilot_sample)` participants (`r sum(pilot_sample$gender == "Female")`
female; `r sum(pilot_sample$gender == "Male")` male;
`r sum(pilot_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(pilot_sample$gender, "Prefer") | is.na(pilot_sample$gender))`
undisclosed gender; mean age = `r round(mean(pilot_sample$age), 2)` years).

### Procedure

We presented participants with six different tasks "that people might perform in
their daily lives". The six tasks were randomly drawn from a larger set of 20 
tasks (see @supptbl-tasks for the full list of tasks). For each task, we asked 
participants the following questions on 7-point Likert scales:

- Is this a social task?
- Does this task require social skills?
- Does this task impact other people?
- How important are the consequences of this task?
- How important is it that effort goes into this task?
- How important is it that others see the effort that goes into this task?

### Statistical analysis

We fitted a Bayesian multivariate multilevel cumulative-link ordinal model to 
the data using the *brms* R package. We modelled each task evaluation as a 
separate response variable and included correlated varying intercepts for
participants and tasks. We used regularising priors for all parameters to impose 
conservatism on parameter estimates (see Supplementary Materials for full model 
specification). The model converged normally ($\hat{R}$ ≤ 1.01).

## Results

We found that participants' responses to all six questions tended to be
positively correlated. For example, tasks rated as more social were also rated 
as requiring more social skills (see @suppfig-tasks-correlations). Estimated 
averages and rankings for the 20 tasks across each of the questions can be found 
in Supplementary Figures [-@suppfig-tasks-social] -- 
[-@suppfig-tasks-extrinsiceffort].

\newpage

# Pilot Study 2

## Methods

### Participants

We conducted a power simulation to determine our target sample size. The 
simulation suggested that a sample size of 150 participants per condition 
(overall *n* = 450 for three conditions) would be required to detect a small 
difference between conditions (Cohen's *d* ≈ 0.20) with above 80% power.

```{r}
study1_sample <-
  tar_read(study1_data) %>%
  group_by(id) %>%
  summarise(
    gender = unique(gender),
    age = unique(age),
    chatgpt_used = unique(chatgpt_used)
    )
```

We recruited a convenience sample of 500 participants from the United Kingdom 
through Prolific. After excluding participants who failed our pre-treatment 
attention check, we were left with a final sample of 
`r nrow(study1_sample)` participants (`r sum(study1_sample$gender == "Female")`
female; `r sum(study1_sample$gender == "Male")` male;
`r sum(study1_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(study1_sample$gender, "Prefer") | is.na(study1_sample$gender))`
undisclosed gender; mean age = `r round(mean(study1_sample$age), 2)` years).
`r round(mean(study1_sample$chatgpt_used == "Yes", na.rm = TRUE) * 100, 0)`% of
these participants reported having used ChatGPT before (see 
@suppfig-chatgpt-pilotstudy2).

### Design

We randomly allocated participants into one of three conditions in a 
between-subjects design: (*i*) the control condition, (*ii*) the AI outsourcing
condition, or (*iii*) the human outsourcing condition. These conditions 
determined how scenarios were presented to participants.

### Procedure

We presented participants with six scenarios (see @suppfig-scenarios-pilotstudy2 
for examples). Each scenario described a person completing a task, such as 
writing computer code or writing a love letter. The six tasks were randomly 
drawn from a larger set of 20 tasks (see @supptbl-tasks for the full list of 
tasks). For each scenario, we told participants:

- *Control condition*: "In order to complete this task, [the person] works on it 
by themselves from start to finish."
- *AI outsourcing condition*: "In order to complete this task, [the person] gets
the AI tool ChatGPT to do it for them."
- *Human outsourcing condition*: "In order to complete this task, [the person]
gets someone else to do it for them."

We then asked participants how well each of the following words described the 
person in the scenario: competent, warm, moral, lazy, and trustworthy.
Participants answered these questions on 7-point Likert scales, ranging from 
"does not describe [the person] well" to "describes [the person] extremely well".

After the six scenarios, we asked participants several questions about the AI 
tool ChatGPT, including their familiarity with ChatGPT, whether they had used 
ChatGPT before, how frequently they used ChatGPT, and how trustworthy they 
thought ChatGPT was (see @suppfig-chatgpt-pilotstudy2).

### Pre-registration

We pre-registered the study on the Open Science Framework
([https://osf.io/khr42](https://osf.io/khr42)).

### Statistical analysis

We fitted Bayesian multivariate multilevel cumulative-link ordinal models to the
data using the *brms* R package. We modelled each character evaluation -- 
competence, warmth, morality, laziness, and trustworthiness -- as a separate 
response variable and included fixed effects for conditions, varying intercepts
for participants, and varying intercepts and slopes for tasks. We used
regularising priors for all parameters to impose conservatism on parameter
estimates (see Supplementary Materials for full model specifications). All
models converged normally ($\hat{R}$ ≤ 1.01).

## Results

We found that people who outsourced tasks to AI or other humans were perceived 
more negatively than people who completed the tasks themselves 
(@suppfig-treatments-pilotstudy2). In particular, people who outsourced were 
perceived as lazier and less competent, with smaller yet detectable differences 
for perceptions of warmth, morality, and trustworthiness 
(@supptbl-treatment-diffs-pilotstudy2). Across all measures, outsourcing to 
other humans was perceived more negatively than outsourcing to AI.

We found that the effects of outsourcing varied across the different tasks, 
especially for perceptions of warmth and morality 
(@suppfig-treatments-by-task-pilotstudy2). For example, people were perceived as 
less warm if they outsourced writing a love letter, but not if they outsourced 
writing computer code. Similarly, people were perceived as less moral if they 
outsourced writing an apology letter to a friend, but not if they outsourced 
writing a dinner recipe. By contrast, the effects of outsourcing on competence, 
laziness, and trustworthiness were more consistent across tasks.

To determine the factors that predict variation across tasks, we incorporated
ratings of tasks from the first pilot study. Participants were asked to rate the 
20 tasks on several features: whether the task is social, requires social skills, 
impacts others, has important consequences, and requires effort. All of these 
features predicted stronger causal effects of outsourcing compared to control
(Supplementary Figures [-@suppfig-interactions-pilotstudy2] and 
[-@suppfig-interaction-pars-pilotstudy2]). In other words, outsourcing to AI or 
other humans is perceived more negatively for tasks that have these features, 
compared to tasks without these features.

\newpage

# Methods for Text Analysis in Study 5

To generate frequency lists for each experimental condition in Study 5, we 
created three documents containing the raw text submissions to the open-ended 
question "In your own words, describe how you feel about Adam and why". Each raw 
text submission was paired with a numbered text ID column. The number of 
submissions was roughly equivalent across conditions: the control condition (N = 
196), the tool outsourcing condition (N = 215), and the full outsourcing 
condition (N = 202).

All text processing was conducted using the Basic Unit-Transposable Text 
Experimentation Resource (BUTTER; Version 0.9.4.1; Boyd, 2019). To prepare the 
data, each CSV file was converted into a folder containing individual text files
-- one per submission -- using two plugins: *Read Text from CSV* (Version 1.0.2) 
and *Save .txt Files to Folder* (Version 1.0.6). The settings for 
*Read Text from CSV* were as follows: file encoding = UTF-8, row identifier = 
ID, text column = Text, CSV delimiter = , and CSV quote = ".

To generate frequency lists, we first loaded the .txt files using the 
*Load .txt Files from Folder* plugin (Version 1.0.4). Tokenization was performed 
using the Twitter-Aware Tokenizer (Version 1.0.2), with the options 
*convert text to lowercase* and *reduce elongation* enabled to minimize 
superficial variation in tokens. We removed filler and function words using the 
*Remove Stop Words* plugin (Version 1.0.31), applying the default English stop 
word list.

Frequency lists were created with the *Frequency List* plugin (Version 1.0.11). 
Settings included: unigram analysis (N = 1), omission of n-grams with frequency
< 5, exclusion of n-grams appearing in fewer than 0.1% of documents, filtering 
collocates by Normalized Pointwise Mutual Information, and removal of collocates 
with metric values < 0.5. Outputs were saved using the *Save Output to CSV* 
plugin (Version 1.0.5). This process was repeated separately for each condition 
folder.

For cross-condition comparison, we used the *Compare Frequencies* plugin 
(Version 1.1.02), retaining most default settings. The only modification was 
disabling the *Skip comparisons with 0 frequency values* option. This plugin 
calculates a range of comparative metrics, including log likelihood (LL), %DIFF, 
Bayes Information Criterion (BIC), relative risk (RRisk), log ratio, and odds 
ratio.

Following previous work (e.g., Rayson & Garside, 2000; Gregson et al., 2022), we 
interpret %DIFF as an indicator of effect size and direction. Frequentist 
statistical significance was determined using log likelihood values, with the 
following thresholds: LL ≥ 3.84 (*p* < .05), LL ≥ 6.63 (*p* < .01), LL ≥ 10.83 
(*p* < .001), and LL ≥ 15.13 (*p* < .0001).

\newpage

# Supplementary Figures

\vspace*{15mm}

::: {#suppfig-tasks-correlations}

```{r tasks-correlations, fig.height=4, fig.width=5}
tar_read(plot_task_cors)
```

Model-estimated task-specific correlations between all six questions. Values are 
posterior median correlations. A positive correlation indicates that tasks that 
are rated highly on one question tend to be rated highly on another question.

:::

\newpage

::: {#suppfig-tasks-social}

```{r tasks-social, fig.height=4, fig.width=6}
tar_read(plot_task_rank_social)
```

Model-estimated means for the question "Is this a social task?" across all 20 
tasks. Grey points represent participant responses to the question, jittered for 
easier viewing. Black points are estimated means from the fitted model, pooling 
over participants. Black points and line ranges represent posterior medians and 
95% credible intervals, respectively.

:::

\newpage

::: {#suppfig-tasks-socialskills}

```{r tasks-socialskills, fig.height=4, fig.width=6}
tar_read(plot_task_rank_socialskills)
```

Model-estimated means for the question "Does this task require social skills?" 
across all 20 tasks. Grey points represent participant responses to the 
question, jittered for easier viewing. Black points are estimated means from the 
fitted model, pooling over participants. Black points and line ranges represent 
posterior medians and 95% credible intervals, respectively.

:::

\newpage

::: {#suppfig-tasks-impactothers}

```{r tasks-impactothers, fig.height=4, fig.width=6}
tar_read(plot_task_rank_impactothers)
```

Model-estimated means for the question "Does this task impact other people?" 
across all 20 tasks. Grey points represent participant responses to the 
question, jittered for easier viewing. Black points are estimated means from the 
fitted model, pooling over participants. Black points and line ranges represent 
posterior medians and 95% credible intervals, respectively.

:::

\newpage

::: {#suppfig-tasks-consequences}

```{r tasks-consequences, fig.height=4, fig.width=6}
tar_read(plot_task_rank_consequences)
```

Model-estimated means for the question "How important are the consequences of 
this task?" across all 20 tasks. Grey points represent participant responses to 
the question, jittered for easier viewing. Black points are estimated means from 
the fitted model, pooling over participants. Black points and line ranges 
represent posterior medians and 95% credible intervals, respectively.

:::

\newpage

::: {#suppfig-tasks-intrinsiceffort}

```{r tasks-intrinsiceffort, fig.height=4, fig.width=6}
tar_read(plot_task_rank_intrinsiceffort)
```

Model-estimated means for the question "How important is it that effort goes 
into this task?" across all 20 tasks. Grey points represent participant 
responses to the question, jittered for easier viewing. Black points are 
estimated means from the fitted model, pooling over participants. Black points 
and line ranges represent posterior medians and 95% credible intervals, 
respectively.

:::

\newpage

::: {#suppfig-tasks-extrinsiceffort}

```{r tasks-extrinsiceffort, fig.height=4, fig.width=6}
tar_read(plot_task_rank_extrinsiceffort)
```

Model-estimated means for the question "How important is it that others see the 
effort that goes into this task?" across all 20 tasks. Grey points represent 
participant responses to the question, jittered for easier viewing. Black points 
are estimated means from the fitted model, pooling over participants. Black 
points and line ranges represent posterior medians and 95% credible intervals, 
respectively.

:::

\newpage

::: {#suppfig-chatgpt-pilotstudy2}

```{r chatgpt-pilotstudy2, fig.height=5, fig.width=6}
tar_read(plot_chatgpt_study1)
```

Responses to the questions about ChatGPT in the second pilot study.

:::

\newpage

::: {#suppfig-scenarios-pilotstudy2}

![](images/scenarios_pilotstudy2.png)

Examples of the scenarios presented to participants in the second pilot study.

:::

\newpage

::: {#suppfig-treatments-pilotstudy2}

```{r treatments-pilotstudy2, fig.height=4, fig.width=6}
tar_read(plot_treatments_study1)
```

Character evaluations in the second pilot study. Participants in the control 
condition, the AI outsourcing condition, and the human outsourcing condition 
evaluated people in the scenarios on (a) competence, (b) warmth, (c) morality,
(d) laziness, and (e) trustworthiness. Coloured points represent participant 
responses to the questions, jittered for easier viewing. Black points are 
estimated marginal means from the fitted model, pooling over participants and 
tasks. Black points and line ranges represent posterior medians and 95% credible 
intervals, respectively.

:::

\newpage

::: {#suppfig-treatments-by-task-pilotstudy2}

```{r treatments-by-task-pilotstudy2, fig.height=5, fig.width=8.5}
tar_read(plot_treatments_tasks_study1)
```

Variation in the effects of outsourcing across tasks in the second pilot study. 
Tasks are ordered from most social (top) to least social (bottom) according to 
ratings from the first pilot study. Point ranges are differences in marginal 
means on a 7-point Likert scale for the AI outsourcing condition (red) and 
the human outsourcing condition (blue) compared to the control condition. Points 
and ranges represent posterior medians and 95% credible intervals, respectively.

:::

\newpage

::: {#suppfig-interactions-pilotstudy2}

```{r interactions-pilotstudy2, fig.height=5, fig.width=6.6}
tar_read(plot_interactions_study1)
```

The impact of task-specific features (e.g., being a social task) on the causal
effects of outsourcing to AI (red) and humans (blue) compared to the control
condition. The y-axis reflects the estimated differences between the
experimental conditions and the control condition (dashed line) on a 7-point 
Likert scale. Lines and shaded areas represent posterior medians and 95%
credible intervals, respectively. The patterns indicate, for example, more 
negative effects of outsourcing on character evaluations for tasks that are 
rated as more social.

:::

\newpage

::: {#suppfig-interaction-pars-pilotstudy2}

```{r interaction-pars-pilotstudy2, fig.height=5.5, fig.width=5.5}
tar_read(plot_interaction_parameters_study1)
```

Interaction parameters from models including task-specific features as
moderators of the causal effects of AI outsourcing (red) and human outsourcing 
(blue) compared to the control condition. Points and line ranges represent 
posterior medians and 95% credible intervals, respectively.

:::

\newpage

# Supplementary Tables

\vspace*{15mm}

::: {#supptbl-tasks tbl-pos="H" layout-ncol="1"}

```{r tasks}
tar_read(table_tasks) %>%
  kable(
    format = "latex",
    escape = FALSE,
    booktabs = TRUE,
    linesep = "", # to prevent space every five rows
    align = "lccccc"
  ) %>%
  kable_styling(font_size = 10)
```

Tasks included in the studies.

:::

\newpage

::: {#supptbl-treatment-diffs-pilotstudy2 tbl-pos="H" layout-ncol="1"}

```{r treatment-diffs-pilotstudy2}
tar_read(table_treatment_diffs_study1) %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    linesep = ""
  ) %>%
  kable_styling(font_size = 9) %>%
  add_header_above(header = c(" " = 1, "Response" = 5))
```

Pairwise contrasts in the second pilot study. Numbers reflect differences in 
marginal means on a 7-point Likert scale, pooling over participants and tasks.
Main numbers are posterior medians, numbers in the square brackets are 95% 
credible intervals.

:::

\newpage

# Supplementary References

Boyd, R. L. (2019). BUTTER: Basic unit-transposable text experimentation 
resource. Available from https://www.butter.tools/

Gregson, R., Piazza, J., & Boyd, R. L. (2022). 'Against the cult of veganism': Unpacking the social psychology and ideology of anti-vegans. *Appetite*, *178*, 106143. doi:10.1016/j.appet.2022.106143

Rayson, P., & Garside, R. (2000). Comparing corpora using frequency profiling. *Proceedings of the Workshop on Comparing Corpora - Volume 9*, 1–6. Presented in Hong Kong. doi:10.3115/1117729.1117730