### 2025-04-28 appendix.qmd
First draft of manuscript

\vspace*{60mm}

# Supplementary Materials

\setcounter{page}{1}
\begin{center}
Negative Perceptions of Outsourcing to Artificial Intelligence \\
Scott Claessens, Pierce Veitch, and Jim Everett \\
School of Psychology, University of Kent
\end{center}
\newpage

# Pilot Study 1

## Methods

### Participants

```{r}
pilot_sample <-
  tar_read(pilot_data) %>%
  group_by(id) %>%
  summarise(
    gender = unique(gender),
    age = unique(age)
    )
```

We recruited a convenience sample of 200 participants from the United Kingdom
through Prolific. After excluding participants who failed our pre-treatment 
attention check, we were left with a final sample of 
`r nrow(pilot_sample)` participants (`r sum(pilot_sample$gender == "Female")`
female; `r sum(pilot_sample$gender == "Male")` male;
`r sum(pilot_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(pilot_sample$gender, "Prefer") | is.na(pilot_sample$gender))`
undisclosed gender; mean age = `r round(mean(pilot_sample$age), 2)` years).

### Procedure

We presented participants with six different tasks "that people might perform in
their daily lives". The six tasks were randomly drawn from a larger set of 20 
tasks (see @supptbl-tasks for the full list of tasks). For each task, we asked 
participants the following questions on 7-point Likert scales:

- Is this a social task?
- Does this task require social skills?
- Does this task impact other people?
- How important are the consequences of this task?
- How important is it that effort goes into this task?
- How important is it that others see the effort that goes into this task?

### Statistical Analysis

We fitted a Bayesian multivariate multilevel cumulative-link ordinal model to 
the data using the *brms* R package. We modelled each task evaluation as a 
separate response variable and included correlated varying intercepts for
participants and tasks. We used regularising priors for all parameters to impose 
conservatism on parameter estimates (see Supplementary Materials for full model 
specification). The model converged normally ($\hat{R}$ ≤ 1.01).

## Results

We found that participants' responses to all six questions tended to be
positively correlated. For example, tasks rated as more social were also rated 
as requiring more social skills (see @suppfig-tasks-correlations). Estimated 
averages and rankings for the 20 tasks across each of the questions can be found 
in Supplementary Figures [-@suppfig-tasks-social] -- 
[-@suppfig-tasks-extrinsiceffort].

\newpage

# Pilot Study 2

## Methods

### Participants

We conducted a power simulation to determine our target sample size. The 
simulation suggested that a sample size of 150 participants per condition 
(overall *n* = 450 for three conditions) would be required to detect a small 
difference between conditions (Cohen's *d* ≈ 0.20) with above 80% power.

```{r}
study1_sample <-
  tar_read(study1_data) %>%
  group_by(id) %>%
  summarise(
    gender = unique(gender),
    age = unique(age),
    chatgpt_used = unique(chatgpt_used)
    )
```

We recruited a convenience sample of 500 participants from the United Kingdom 
through Prolific. After excluding participants who failed our pre-treatment 
attention check, we were left with a final sample of 
`r nrow(study1_sample)` participants (`r sum(study1_sample$gender == "Female")`
female; `r sum(study1_sample$gender == "Male")` male;
`r sum(study1_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(study1_sample$gender, "Prefer") | is.na(study1_sample$gender))`
undisclosed gender; mean age = `r round(mean(study1_sample$age), 2)` years).
`r round(mean(study1_sample$chatgpt_used == "Yes", na.rm = TRUE) * 100, 0)`% of
these participants reported having used ChatGPT before (see 
@suppfig-chatgpt-pilotstudy2).

### Design

We randomly allocated participants into one of three conditions in a 
between-subjects design: (*i*) the control condition, (*ii*) the AI outsourcing
condition, or (*iii*) the human outsourcing condition. These conditions 
determined how scenarios were presented to participants.

### Procedure

We presented participants with six scenarios. Each scenario described a person 
completing a task, such as writing computer code or writing a love letter. The 
six tasks were randomly drawn from a larger set of 20 tasks (see @supptbl-tasks
for the full list of tasks). For each scenario, we told participants:

- *Control condition*: "In order to complete this task, [the person] works on it 
by themselves from start to finish."
- *AI outsourcing condition*: "In order to complete this task, [the person] gets
the AI tool ChatGPT to do it for them."
- *Human outsourcing condition*: "In order to complete this task, [the person]
gets someone else to do it for them."

We then asked participants how well each of the following words described the 
person in the scenario: competent, warm, moral, lazy, and trustworthy.
Participants answered these questions on 7-point Likert scales, ranging from 
"does not describe [the person] well" to "describes [the person] extremely well".

After the six scenarios, we asked participants several questions about the AI 
tool ChatGPT, including their familiarity with ChatGPT, whether they had used 
ChatGPT before, how frequently they used ChatGPT, and how trustworthy they 
thought ChatGPT was (see @suppfig-chatgpt-pilotstudy2).

### Pre-registration

We pre-registered the study on the Open Science Framework
(<https://osf.io/khr42>).

### Statistical Analysis

We fitted Bayesian multivariate multilevel cumulative-link ordinal models to the
data using the *brms* R package. We modelled each character evaluation -- 
competence, warmth, morality, laziness, and trustworthiness -- as a separate 
response variable and included fixed effects for conditions, varying intercepts
for participants, and varying intercepts and slopes for tasks. We used
regularising priors for all parameters to impose conservatism on parameter
estimates (see Supplementary Materials for full model specifications). All
models converged normally ($\hat{R}$ ≤ 1.01).

## Results

We found that people who outsourced tasks to AI or other humans were perceived 
more negatively than people who completed the tasks themselves 
(@suppfig-treatments-pilotstudy2). In particular, people who outsourced were 
perceived as lazier and less competent, with smaller yet detectable differences 
for perceptions of warmth, morality, and trustworthiness 
(@supptbl-treatment-diffs-pilotstudy2). Across all measures, outsourcing to 
other humans was perceived more negatively than outsourcing to AI.

We found that the effects of outsourcing varied across the different tasks, 
especially for perceptions of warmth and morality 
(@suppfig-treatments-by-task-pilotstudy2). For example, people were perceived as 
less warm if they outsourced writing a love letter, but not if they outsourced 
writing computer code. Similarly, people were perceived as less moral if they 
outsourced writing an apology letter to a friend, but not if they outsourced 
writing a dinner recipe. By contrast, the effects of outsourcing on competence, 
laziness, and trustworthiness were more consistent across tasks.

To determine the factors that predict variation across tasks, we incorporated
ratings of tasks from the first pilot study. Participants were asked to rate the 
20 tasks on several features: whether the task is social, requires social skills, 
impacts others, has important consequences, and requires effort. All of these 
features predicted stronger causal effects of outsourcing compared to control
(Supplementary Figures [-@suppfig-interactions-pilotstudy2] and 
[-@suppfig-interaction-pars-pilotstudy2]). In other words, outsourcing to AI or 
other humans is perceived more negatively for tasks that have these features, 
compared to tasks without these features.

\newpage

# Vignette Wording in Study 5

:::{.NoIndent}
We presented participants in Study 5 with the following vignette text:
:::

> Adam has been dating his partner for almost a year, and Valentine's Day is 
coming up. He knows that many people exchange a card on Valentine's Day 
containing a love letter to their partner, and he decides to send a love letter
to his partner too.
>
> This year, Adam has been closely following developments in technology and has
read of people using AI tools like ChatGPT for things like this, either using it
to help with writing or getting AI to do the task completely.

:::{.NoIndent}
This was followed by the manipulation text (see main text). Participants were 
then presented with the love letter that Adam ostensibly wrote, which was held 
constant across conditions:
:::

> Happy Valentine's Day, my love.
>
> I don't think I tell you enough just how much you mean to me. Being with you 
feels like breathing a little easier, like the world is a bit softer just 
because you're in it. You make the everyday feel special, and somehow you always
know how to calm my nerves or make me laugh at just the right moment. I feel
like myself with you -- maybe even a better version of myself -- and that's such
a rare and beautiful thing.
>
> I'm so grateful for you -- for the way you listen, the way you love, the way
you show up, even in the small ways. I hope you know that no matter what, I'm
always in your corner. I can't wait to keep making memories together, whether
we're off on some adventure or just curled up on the couch. I love you more than
I can really put into words, but I promise I'll spend every day trying.
>
> Yours,
>
> Adam

\newpage

# Methods for Text Analysis in Study 5

To generate frequency lists for each experimental condition in Study 5, we 
created three documents containing the raw text submissions to the open-ended 
question "In your own words, describe how you feel about Adam and why". Each raw 
text submission was paired with a numbered text ID column. The number of 
submissions was roughly equivalent across conditions: the control condition (N = 
196), the tool outsourcing condition (N = 215), and the full outsourcing 
condition (N = 202).

All text processing was conducted using the Basic Unit-Transposable Text 
Experimentation Resource (BUTTER; Version 0.9.4.1; Boyd, 2019). To prepare the 
data, each CSV file was converted into a folder containing individual text files
-- one per submission -- using two plugins: *Read Text from CSV* (Version 1.0.2) 
and *Save .txt Files to Folder* (Version 1.0.6). The settings for 
*Read Text from CSV* were as follows: file encoding = UTF-8, row identifier = 
ID, text column = Text, CSV delimiter = , and CSV quote = ".

To generate frequency lists, we first loaded the .txt files using the 
*Load .txt Files from Folder* plugin (Version 1.0.4). Tokenization was performed 
using the Twitter-Aware Tokenizer (Version 1.0.2), with the options 
*convert text to lowercase* and *reduce elongation* enabled to minimize 
superficial variation in tokens. We removed filler and function words using the 
*Remove Stop Words* plugin (Version 1.0.31), applying the default English stop 
word list.

Frequency lists were created with the *Frequency List* plugin (Version 1.0.11). 
Settings included: unigram analysis (N = 1), omission of n-grams with frequency
< 5, exclusion of n-grams appearing in fewer than 0.1% of documents, filtering 
collocates by Normalized Pointwise Mutual Information (NPMI), and removal of 
collocates with metric values < 0.5. Outputs were saved using the 
*Save Output to CSV* plugin (Version 1.0.5). This process was repeated 
separately for each condition folder.

For cross-condition comparison, we used the *Compare Frequencies* plugin 
(Version 1.1.02), retaining most default settings. The only modification was 
disabling the *Skip comparisons with 0 frequency values* option. This plugin 
calculates a range of comparative metrics, including log likelihood (LL), %DIFF, 
Bayes Information Criterion (BIC), relative risk (RRisk), log ratio, and odds 
ratio.

Following previous work (e.g., Rayson & Garside, 2000; Gregson et al., 2022), we 
interpret %DIFF as an indicator of effect size and direction. Frequentist 
statistical significance was determined using log likelihood values, with the 
following thresholds: LL ≥ 3.84 (*p* < .05), LL ≥ 6.63 (*p* < .01), LL ≥ 10.83 
(*p* < .001), and LL ≥ 15.13 (*p* < .0001).

\newpage

# Supplementary Figures

\vspace*{15mm}

::: {#suppfig-chatgpt-study1 suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot_chatgpt_study2)
```

Responses to the questions about ChatGPT in Study 1.

:::

\newpage

::: {#suppfig-interactions-study1 suppfig-pos="H"}

```{r, fig.height=6, fig.width=7}
tar_read(plot_interactions_study2)
```

The impact of task-specific features (e.g., being a social task) on the causal
effects of outsourcing to AI compared to the control condition in Study 1. The 
y-axis reflects the estimated differences between the experimental conditions 
and the control condition (dashed line) on a 7-point Likert scale. Lines and 
shaded areas represent posterior medians and 95% credible intervals, 
respectively. The patterns indicate, for example, more negative effects of 
outsourcing on character evaluations for tasks that are rated as more social.

:::

\newpage

::: {#suppfig-interaction-pars-study1 suppfig-pos="H"}

```{r, fig.height=6, fig.width=6}
tar_read(plot_interaction_parameters_study2)
```

Interaction parameters from models including task-specific features as
moderators of the causal effects of AI outsourcing compared to the control 
condition in Study 1. Points and line ranges represent posterior medians and 95% 
credible intervals, respectively.

:::

\newpage

::: {#suppfig-chatgpt-study2 suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot_chatgpt_study3)
```

Responses to the questions about ChatGPT in Study 2.

:::

\newpage

::: {#suppfig-treatments-person-study2 suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_treatments_person_study3)
```

Character evaluations in Study 2. Participants in the control condition, the 
tool outsourcing condition, and the full outsourcing condition evaluated the 
"other participant" on (a) competence, (b) warmth, (c) morality, (d) laziness, 
and (e) trustworthiness. Jittered points represent participant responses to the
questions, split by whether the writing task was a non-social task (red) or a 
social task (blue). Point ranges are estimated marginal means from the fitted 
model, pooling over essay answers. Points and line ranges represent posterior 
medians and 95% credible intervals, respectively.

:::

\newpage

::: {#suppfig-chatgpt-study3 suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot_chatgpt_study4)
```

Responses to the questions about ChatGPT in Study 3.

:::

\newpage

::: {#suppfig-treatments-tasks-study3 suppfig-pos="H"}

```{r, fig.height=5, fig.width=8.5}
tar_read(plot_treatments_tasks_study4)
```

Variation in the effects of outsourcing across tasks in Study 3. Tasks are 
ordered from most social (top) to least social (bottom) according to ratings 
from a pilot study. Point ranges are differences in marginal means on a 7-point
Likert scale for the low effort conditions (red) and high effort conditions 
(blue) compared to the control condition. Upper panels refer to the standard LLM
conditions, and lower panels refer to the personalised LLM conditions. Points 
and ranges represent posterior medians and 95% credible intervals, respectively.

:::

\newpage

::: {#suppfig-interactions-study3 suppfig-pos="H"}

```{r, fig.height=6, fig.width=7}
tar_read(plot_interactions_study4)
```

The impact of task-specific features (e.g., being a social task) on the causal
effects of outsourcing to AI compared to the control condition in Study 3. The 
y-axis reflects the estimated differences between the experimental conditions 
and the control condition (dashed line) on a 7-point Likert scale. Lines and 
shaded areas represent posterior medians and 95% credible intervals, 
respectively. The patterns indicate, for example, more negative effects of 
outsourcing on character evaluations for tasks that are rated as more social.

:::

\newpage

::: {#suppfig-interaction-pars-study3 suppfig-pos="H"}

```{r, fig.height=6, fig.width=6}
tar_read(plot_interaction_parameters_study4)
```

Interaction parameters from models including task-specific features as
moderators of the causal effects of AI outsourcing compared to the control 
condition in Study 3. Points and line ranges represent posterior medians and 95% 
credible intervals, respectively.

:::

\newpage

::: {#suppfig-chatgpt-study4 suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot_chatgpt_study5)
```

Responses to the questions about ChatGPT in Study 4.

:::

\newpage

::: {#suppfig-treatments-tasks-study4 suppfig-pos="H"}

```{r, fig.height=4, fig.width=8.5}
tar_read(plot_treatments_tasks_study5)
```

Variation in the effects of outsourcing across tasks in Study 4. Tasks are 
ordered from most social (top) to least social (bottom) according to ratings 
from a pilot study. Point ranges are differences in marginal means on a 7-point
Likert scale for the bad reason conditions (red) and good reason conditions 
(blue) compared to the control condition. Upper panels refer to the tool 
outsourcing conditions, and lower panels refer to the full outsourcing 
conditions. Points and ranges represent posterior medians and 95% credible 
intervals, respectively.

:::

\newpage

::: {#suppfig-interactions-study4 suppfig-pos="H"}

```{r, fig.height=6, fig.width=7}
tar_read(plot_interactions_study5)
```

The impact of task-specific features (e.g., being a social task) on the causal
effects of outsourcing to AI compared to the control condition in Study 4. The 
y-axis reflects the estimated differences between the experimental conditions 
and the control condition (dashed line) on a 7-point Likert scale. Lines and 
shaded areas represent posterior medians and 95% credible intervals, 
respectively. The patterns indicate, for example, more negative effects of 
outsourcing on character evaluations for tasks that are rated as more social.

:::

\newpage

::: {#suppfig-interaction-pars-study4 suppfig-pos="H"}

```{r, fig.height=6, fig.width=6}
tar_read(plot_interaction_parameters_study5)
```

Interaction parameters from models including task-specific features as
moderators of the causal effects of AI outsourcing compared to the control 
condition in Study 4. Points and line ranges represent posterior medians and 95% 
credible intervals, respectively.

:::

\newpage

::: {#suppfig-chatgpt-study5 suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot_chatgpt_study6)
```

Responses to the questions about ChatGPT in Study 5.

:::

\newpage

::: {#suppfig-tasks-correlations suppfig-pos="H"}

```{r, fig.height=4, fig.width=5}
tar_read(plot_task_cors)
```

Model-estimated task-specific correlations between all six questions in the 
first pilot study. Values are posterior median correlations. A positive 
correlation indicates that tasks that are rated highly on one question tend to 
be rated highly on another question.

:::

\newpage

::: {#suppfig-tasks-social suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_task_rank_social)
```

Model-estimated means for the question "Is this a social task?" across all 20 
tasks in the first pilot study. Grey points represent participant responses to 
the question, jittered for easier viewing. Black points are estimated means from
the fitted model, pooling over participants. Black points and line ranges 
represent posterior medians and 95% credible intervals, respectively.

:::

\newpage

::: {#suppfig-tasks-socialskills suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_task_rank_socialskills)
```

Model-estimated means for the question "Does this task require social skills?" 
across all 20 tasks in the first pilot study. Grey points represent participant 
responses to the question, jittered for easier viewing. Black points are 
estimated means from the fitted model, pooling over participants. Black points 
and line ranges represent posterior medians and 95% credible intervals, 
respectively.

:::

\newpage

::: {#suppfig-tasks-impactothers suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_task_rank_impactothers)
```

Model-estimated means for the question "Does this task impact other people?" 
across all 20 tasks in the first pilot study. Grey points represent participant 
responses to the question, jittered for easier viewing. Black points are 
estimated means from the fitted model, pooling over participants. Black points 
and line ranges represent posterior medians and 95% credible intervals,
respectively.

:::

\newpage

::: {#suppfig-tasks-consequences suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_task_rank_consequences)
```

Model-estimated means for the question "How important are the consequences of 
this task?" across all 20 tasks in the first pilot study. Grey points represent 
participant responses to the question, jittered for easier viewing. Black points
are estimated means from the fitted model, pooling over participants. Black
points and line ranges represent posterior medians and 95% credible intervals, 
respectively.

:::

\newpage

::: {#suppfig-tasks-intrinsiceffort suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_task_rank_intrinsiceffort)
```

Model-estimated means for the question "How important is it that effort goes 
into this task?" across all 20 tasks in the first pilot study. Grey points 
represent participant responses to the question, jittered for easier viewing.
Black points are estimated means from the fitted model, pooling over
participants. Black points and line ranges represent posterior medians and 95% 
credible intervals, respectively.

:::

\newpage

::: {#suppfig-tasks-extrinsiceffort suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_task_rank_extrinsiceffort)
```

Model-estimated means for the question "How important is it that others see the 
effort that goes into this task?" across all 20 tasks in the first pilot study.
Grey points represent participant responses to the question, jittered for easier 
viewing. Black points are estimated means from the fitted model, pooling over 
participants. Black points and line ranges represent posterior medians and 95% 
credible intervals, respectively.

:::

\newpage

::: {#suppfig-chatgpt-pilotstudy2 suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot_chatgpt_study1)
```

Responses to the questions about ChatGPT in the second pilot study.

:::

\newpage

::: {#suppfig-treatments-pilotstudy2 suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_treatments_study1)
```

Character evaluations in the second pilot study. Participants in the control 
condition, the AI outsourcing condition, and the human outsourcing condition 
evaluated people in the scenarios on (a) competence, (b) warmth, (c) morality,
(d) laziness, and (e) trustworthiness. Coloured points represent participant 
responses to the questions, jittered for easier viewing. Black points are 
estimated marginal means from the fitted model, pooling over participants and 
tasks. Black points and line ranges represent posterior medians and 95% credible 
intervals, respectively.

:::

\newpage

::: {#suppfig-treatments-by-task-pilotstudy2 suppfig-pos="H"}

```{r, fig.height=5, fig.width=8.5}
tar_read(plot_treatments_tasks_study1)
```

Variation in the effects of outsourcing across tasks in the second pilot study. 
Tasks are ordered from most social (top) to least social (bottom) according to 
ratings from the first pilot study. Point ranges are differences in marginal 
means on a 7-point Likert scale for the AI outsourcing condition (red) and 
the human outsourcing condition (blue) compared to the control condition. Points 
and ranges represent posterior medians and 95% credible intervals, respectively.

:::

\newpage

::: {#suppfig-interactions-pilotstudy2 suppfig-pos="H"}

```{r, fig.height=5, fig.width=6.6}
tar_read(plot_interactions_study1)
```

The impact of task-specific features (e.g., being a social task) on the causal
effects of outsourcing to AI (red) and humans (blue) compared to the control
condition in the second pilot study. The y-axis reflects the estimated 
differences between the experimental conditions and the control condition 
(dashed line) on a 7-point Likert scale. Lines and shaded areas represent 
posterior medians and 95% credible intervals, respectively. The patterns 
indicate, for example, more negative effects of outsourcing on character 
evaluations for tasks that are rated as more social.

:::

\newpage

::: {#suppfig-interaction-pars-pilotstudy2 suppfig-pos="H"}

```{r, fig.height=5.5, fig.width=5.5}
tar_read(plot_interaction_parameters_study1)
```

Interaction parameters from models including task-specific features as
moderators of the causal effects of AI outsourcing (red) and human outsourcing 
(blue) compared to the control condition in the second pilot study. Points and 
line ranges represent posterior medians and 95% credible intervals,
respectively.

:::

\newpage

# Supplementary Tables

::: {#supptbl-tasks supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_tasks) %>%
  kable(
    format = "latex",
    escape = FALSE,
    booktabs = TRUE,
    linesep = "",
    align = "lccccc"
  ) %>%
  kable_styling(font_size = 10)
```

Tasks included in the studies.

:::

\newpage

::: {#supptbl-essay-answers-social-study2 supptbl-pos="H" layout-ncol="1"}

```{r}
table <- 
  tar_read(table_essay_answers_study3) %>%
  slice(1:3) %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    linesep = ""
  ) %>%
  kable_styling(font_size = 10) %>%
  column_spec(1, width = "2cm") %>%
  column_spec(2, width = "14cm")

# table single spaced
paste(
  "\\begingroup",
  "\\linespread{1}\\selectfont",
  table,
  "\\endgroup",
  sep = "\n"
  ) %>%
  knitr::asis_output()
```

Pre-generated essay answers to the social prompt in Study 2.

:::

\newpage

::: {#supptbl-essay-answers-nonsocial-study2 supptbl-pos="H" layout-ncol="1"}

```{r}
table <- 
  tar_read(table_essay_answers_study3) %>%
  slice(4:6) %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    linesep = ""
  ) %>%
  kable_styling(font_size = 10) %>%
  column_spec(1, width = "2cm") %>%
  column_spec(2, width = "14cm")

# table single spaced
paste(
  "\\begingroup",
  "\\linespread{1}\\selectfont",
  table,
  "\\endgroup",
  sep = "\n"
  ) %>%
  knitr::asis_output()
```

Pre-generated essay answers to the non-social prompt in Study 2.

:::

\newpage

::: {#supptbl-essay-comprehension-study2 supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_essay_comprehension_study3) %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    linesep = ""
  ) %>%
  kable_styling(font_size = 9)
```

Reading times and comprehension rates for the essay answers in Study 2. Expected
reading times were calculated based on an estimated reading speed of 275 words
per minute. Comprehension rates are the percentage of participants who answered
the comprehension question correctly.

:::

\newpage

::: {#supptbl-manipulation-check-study2 supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_manipulation_check_study3) %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    linesep = ""
  ) %>%
  kable_styling(font_size = 11)
```

Percentage of participants in Study 2 who passed the manipulation check and 
reported that they believed the manipulation, split by condition.

:::

\newpage

::: {#supptbl-treatment-diffs-person-study2 supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_treatment_diffs_effects_study3) %>%
  dplyr::select(c(1, 7:11)) %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    linesep = ""
  ) %>%
  kable_styling(font_size = 7.5) %>%
  add_header_above(header = c(" " = 1, "Response" = 5)) %>%
  pack_rows(
    index = c(
      "Effect of outsourcing type" = 6,
      "Effect of task type" = 3,
      "Interaction effect" = 3
      )
    ) %>%
  pack_rows("\u00A0\u00A0Task type = Social", 1, 3, indent = FALSE) %>%
  pack_rows("\u00A0\u00A0Task type = Non-social", 4, 6, indent = FALSE)
```

Pairwise contrasts for character evaluations in Study 2. Numbers reflect 
differences in marginal means on a 7-point Likert scale. Estimates are pooled 
over essay answers. The bottom rows represent the interactions between 
outsourcing type and task type. Main numbers are posterior medians, numbers in 
the square brackets are 95% credible intervals.

:::

\newpage

::: {.landscape}

::: {#supptbl-text-analysis-study5 supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_text_analysis_study6) %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    linesep = ""
  ) %>%
  kable_styling(font_size = 8)
```

Pairwise comparisons of word frequencies between conditions. LL = log 
likelihood; %DIFF =

:::

:::

\newpage

::: {#supptbl-treatment-diffs-pilotstudy2 supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_treatment_diffs_study1) %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    linesep = ""
  ) %>%
  kable_styling(font_size = 9) %>%
  add_header_above(header = c(" " = 1, "Response" = 5))
```

Pairwise contrasts in the second pilot study. Numbers reflect differences in 
marginal means on a 7-point Likert scale, pooling over participants and tasks. 
Main numbers are posterior medians, numbers in the square brackets are 95% 
credible intervals.

:::

\newpage

# Supplementary References

\setlength{\parindent}{-20pt}
\setlength{\leftskip}{20pt}

Boyd, R. L. (2019). BUTTER: Basic unit-transposable text experimentation 
resource. Available from <https://www.butter.tools/>

Gregson, R., Piazza, J., & Boyd, R. L. (2022). 'Against the cult of veganism': 
Unpacking the social psychology and ideology of anti-vegans. *Appetite*, *178*,
106143. doi:[10.1016/j.appet.2022.106143](https://doi.org/10.1016/j.appet.2022.106143)

Rayson, P., & Garside, R. (2000). Comparing corpora using frequency profiling.
*Proceedings of the Workshop on Comparing Corpora - Volume 9*, 1–6. Presented in
Hong Kong. doi:[10.3115/1117729.1117730](https://doi.org/10.3115/1117729.1117730)