### 2025-06-23 appendix.qmd
Recreate manuscript as docx and submit

# Supplementary Materials

# Pilot Study

## Methods

### Participants

```{r}
pilot_sample <-
  tar_read(pilot_data) %>%
  group_by(id) %>%
  summarise(
    gender = unique(gender),
    age = unique(age)
    )
```

We recruited a convenience sample of 200 participants from the United Kingdom
through Prolific. After excluding participants who failed our pre-treatment
attention check, we were left with a final sample of
`r nrow(pilot_sample)` participants (`r sum(pilot_sample$gender == "Female")`
female; `r sum(pilot_sample$gender == "Male")` male;
`r sum(pilot_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(pilot_sample$gender, "Prefer") | is.na(pilot_sample$gender))`
undisclosed gender; mean age = `r round(mean(pilot_sample$age), 2)` years).

### Procedure

We presented participants with six different tasks "that people might perform in
their daily lives". The six tasks were randomly drawn from a larger set of 20
tasks (see @supptbl-tasks for the full list of tasks). For each task, we asked
participants the following questions on 7-point Likert scales:

- Is this a social task?
- Does this task require social skills?
- Does this task impact other people?
- How important are the consequences of this task?
- How important is it that effort goes into this task?
- How important is it that others see the effort that goes into this task?

<br>

### Statistical Analysis

We fitted a Bayesian multivariate multilevel cumulative-link ordinal model to
the data using the *brms* R package. We modelled each task evaluation as a
separate response variable and included correlated varying intercepts for
participants and tasks. We used regularizing priors for all parameters to impose
conservatism on parameter estimates (see Supplementary Materials for full model
specification). The model converged normally ($\hat{R}$ ≤ 1.01).

## Results

We found that participants' responses to all six questions tended to be
positively correlated. For example, tasks rated as more social were also rated
as requiring more social skills (see @suppfig-tasks-correlations). Estimated
averages and rankings for the 20 tasks across each of the questions can be found
in Supplementary Figures [-@suppfig-tasks-social] --
[-@suppfig-tasks-extrinsiceffort].

# Vignette Wording in Study 6

:::{.NoIndent}
We presented participants in Study 6 with the following vignette text:
:::

> Adam has been dating his partner for almost a year, and Valentine's Day is
coming up. He knows that many people exchange a card on Valentine's Day
containing a love letter to their partner, and he decides to send a love letter
to his partner too.
>
> This year, Adam has been closely following developments in technology and has
read of people using AI tools like ChatGPT for things like this, either using it
to help with writing or getting AI to do the task completely.

:::{.NoIndent}
This was followed by the manipulation text (see main text). Participants were
then presented with the love letter that Adam ostensibly wrote, which was held
constant across conditions:
:::

> Happy Valentine's Day, my love.
>
> I don't think I tell you enough just how much you mean to me. Being with you
feels like breathing a little easier, like the world is a bit softer just
because you're in it. You make the everyday feel special, and somehow you always
know how to calm my nerves or make me laugh at just the right moment. I feel
like myself with you -- maybe even a better version of myself -- and that's such
a rare and beautiful thing.
>
> I'm so grateful for you -- for the way you listen, the way you love, the way
you show up, even in the small ways. I hope you know that no matter what, I'm
always in your corner. I can't wait to keep making memories together, whether
we're off on some adventure or just curled up on the couch. I love you more than
I can really put into words, but I promise I'll spend every day trying.
>
> Yours,
>
> Adam

# Methods for Text Analysis in Study 6

To generate frequency lists for each experimental condition in Study 6, we
created three documents containing the raw text submissions to the open-ended
question "In your own words, describe how you feel about Adam and why". Each raw
text submission was paired with a numbered text ID column. The number of
submissions was roughly equivalent across conditions: the control condition (N =
196), the tool outsourcing condition (N = 215), and the full outsourcing
condition (N = 202).

All text processing was conducted using the Basic Unit-Transposable Text
Experimentation Resource (BUTTER; Version 0.9.4.1; Boyd, 2019). To prepare the
data, each CSV file was converted into a folder containing individual text files
-- one per submission -- using two plugins: *Read Text from CSV* (Version 1.0.2)
and *Save .txt Files to Folder* (Version 1.0.6). The settings for
*Read Text from CSV* were as follows: file encoding = UTF-8, row identifier =
ID, text column = Text, CSV delimiter = , and CSV quote = ".

To generate frequency lists, we first loaded the .txt files using the
*Load .txt Files from Folder* plugin (Version 1.0.4). Tokenization was performed
using the Twitter-Aware Tokenizer (Version 1.0.2), with the options
*convert text to lowercase* and *reduce elongation* enabled to minimize
superficial variation in tokens. We removed filler and function words using the
*Remove Stop Words* plugin (Version 1.0.31), applying the default English stop
word list.

Frequency lists were created with the *Frequency List* plugin (Version 1.0.11).
Settings included: unigram analysis (N = 1), omission of n-grams with frequency
< 5, exclusion of n-grams appearing in fewer than 0.1% of documents, filtering
collocates by Normalized Pointwise Mutual Information (NPMI), and removal of
collocates with metric values < 0.5. Outputs were saved using the
*Save Output to CSV* plugin (Version 1.0.5). This process was repeated
separately for each condition folder.

For cross-condition comparison, we used the *Compare Frequencies* plugin
(Version 1.1.02), retaining most default settings. The only modification was
disabling the *Skip comparisons with 0 frequency values* option. This plugin
calculates a range of comparative metrics, including log likelihood (LL), %DIFF,
Bayes Information Criterion (BIC), relative risk (RRisk), log ratio, and odds
ratio.

Following previous work (e.g., Rayson & Garside, 2000; Gregson et al., 2022), we
interpret %DIFF as an indicator of effect size and direction. Frequentist
statistical significance was determined using log likelihood values, with the
following thresholds: LL ≥ 3.84 (*p* < .05), LL ≥ 6.63 (*p* < .01), LL ≥ 10.83
(*p* < .001), and LL ≥ 15.13 (*p* < .0001).

\newpage

# Supplementary Figures

\vspace*{15mm}

::: {#suppfig-interactions-study1 suppfig-pos="H"}

```{r, fig.height=5, fig.width=6.6}
tar_read(plot_interactions_study1)
```

The impact of task-specific features (e.g., being a social task) on the causal
effects of outsourcing to AI (red) and humans (blue) compared to the control
condition in Study 1. The y-axis reflects the estimated differences between the 
experimental conditions and the control condition (dashed line) on a 7-point 
Likert scale. Lines and shaded areas represent posterior medians and 95% 
credible intervals, respectively. The patterns indicate, for example, more 
negative effects of outsourcing on character evaluations for tasks that are 
rated as more social.

:::

\newpage

::: {#suppfig-interaction-pars-study1 suppfig-pos="H"}

```{r, fig.height=5.5, fig.width=5.5}
tar_read(plot_interaction_parameters_study1)
```

Interaction parameters from models including task-specific features as
moderators of the causal effects of AI outsourcing (red) and human outsourcing
(blue) compared to the control condition in Study 1. Points and line ranges 
represent posterior medians and 95% credible intervals, respectively.

:::

\newpage

::: {#suppfig-interactions-study2 suppfig-pos="H"}

```{r, fig.height=6, fig.width=7}
tar_read(plot_interactions_study2)
```

The impact of task-specific features (e.g., being a social task) on the causal
effects of outsourcing to AI compared to the control condition in Study 2. The
y-axis reflects the estimated differences between the experimental conditions
and the control condition (dashed line) on a 7-point Likert scale. Lines and
shaded areas represent posterior medians and 95% credible intervals,
respectively. The patterns indicate, for example, more negative effects of
outsourcing on character evaluations for tasks that are rated as more social.

:::

\newpage

::: {#suppfig-interaction-pars-study2 suppfig-pos="H"}

```{r, fig.height=6, fig.width=6}
tar_read(plot_interaction_parameters_study2)
```

Interaction parameters from models including task-specific features as
moderators of the causal effects of AI outsourcing compared to the control
condition in Study 2. Points and line ranges represent posterior medians and 95%
credible intervals, respectively.

:::

\newpage

::: {#suppfig-treatments-person-study3 suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_treatments_person_study3)
```

Character evaluations in Study 3. Participants in the control condition, the
tool outsourcing condition, and the full outsourcing condition evaluated the
"other participant" on (a) competence, (b) warmth, (c) morality, (d) laziness,
and (e) trustworthiness. Jittered points represent participant responses to the
questions, split by whether the writing task was a non-social task (red) or a
social task (blue). Point ranges are estimated marginal means from the fitted
model, pooling over essay answers. Points and line ranges represent posterior
medians and 95% credible intervals, respectively.

:::

\newpage

::: {#suppfig-interactions-study4 suppfig-pos="H"}

```{r, fig.height=6, fig.width=7}
tar_read(plot_interactions_study4)
```

The impact of task-specific features (e.g., being a social task) on the causal
effects of outsourcing to AI compared to the control condition in Study 4. The
y-axis reflects the estimated differences between the experimental conditions
and the control condition (dashed line) on a 7-point Likert scale. Lines and
shaded areas represent posterior medians and 95% credible intervals,
respectively. The patterns indicate, for example, more negative effects of
outsourcing on character evaluations for tasks that are rated as more social.

:::

\newpage

::: {#suppfig-interaction-pars-study4 suppfig-pos="H"}

```{r, fig.height=6, fig.width=6}
tar_read(plot_interaction_parameters_study4)
```

Interaction parameters from models including task-specific features as
moderators of the causal effects of AI outsourcing compared to the control
condition in Study 4. Points and line ranges represent posterior medians and 95%
credible intervals, respectively.

:::

\newpage

::: {#suppfig-interactions-study5 suppfig-pos="H"}

```{r, fig.height=6, fig.width=7}
tar_read(plot_interactions_study5)
```

The impact of task-specific features (e.g., being a social task) on the causal
effects of outsourcing to AI compared to the control condition in Study 5. The
y-axis reflects the estimated differences between the experimental conditions
and the control condition (dashed line) on a 7-point Likert scale. Lines and
shaded areas represent posterior medians and 95% credible intervals,
respectively. The patterns indicate, for example, more negative effects of
outsourcing on character evaluations for tasks that are rated as more social.

:::

\newpage

::: {#suppfig-interaction-pars-study5 suppfig-pos="H"}

```{r, fig.height=6, fig.width=6}
tar_read(plot_interaction_parameters_study5)
```

Interaction parameters from models including task-specific features as
moderators of the causal effects of AI outsourcing compared to the control
condition in Study 5. Points and line ranges represent posterior medians and 95%
credible intervals, respectively.

:::

\newpage

::: {.landscape}

::: {#suppfig-reasons-tasks-study5 suppfig-pos="H"}

```{r, fig.height=3.5, fig.width=8.5}
tar_read(plot_reasons_tasks_study5)
```

Variation in the effect of reasons across tasks in Study 5. Tasks are ordered
from most social (top) to least social (bottom) according to ratings from a
pilot study. Point ranges are differences in marginal means on a 7-point Likert
scale between the "bad reason" and "good reason" conditions, split by
outsourcing type. Points and ranges represent posterior medians and 95% credible
intervals, respectively.

:::

:::

\newpage

::: {#suppfig-tasks-correlations suppfig-pos="H"}

```{r, fig.height=4, fig.width=5}
tar_read(plot_task_cors)
```

Model-estimated task-specific correlations between all six questions in the
pilot study. Values are posterior median correlations. A positive correlation 
indicates that tasks that are rated highly on one question tend to be rated 
highly on another question.

:::

\newpage

::: {#suppfig-tasks-social suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_task_rank_social)
```

Model-estimated means for the question "Is this a social task?" across all 20
tasks in the pilot study. Grey points represent participant responses to
the question, jittered for easier viewing. Black points are estimated means from
the fitted model, pooling over participants. Black points and line ranges
represent posterior medians and 95% credible intervals, respectively.

:::

\newpage

::: {#suppfig-tasks-socialskills suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_task_rank_socialskills)
```

Model-estimated means for the question "Does this task require social skills?"
across all 20 tasks in the pilot study. Grey points represent participant
responses to the question, jittered for easier viewing. Black points are
estimated means from the fitted model, pooling over participants. Black points
and line ranges represent posterior medians and 95% credible intervals,
respectively.

:::

\newpage

::: {#suppfig-tasks-impactothers suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_task_rank_impactothers)
```

Model-estimated means for the question "Does this task impact other people?"
across all 20 tasks in the pilot study. Grey points represent participant
responses to the question, jittered for easier viewing. Black points are
estimated means from the fitted model, pooling over participants. Black points
and line ranges represent posterior medians and 95% credible intervals,
respectively.

:::

\newpage

::: {#suppfig-tasks-consequences suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_task_rank_consequences)
```

Model-estimated means for the question "How important are the consequences of
this task?" across all 20 tasks in the pilot study. Grey points represent
participant responses to the question, jittered for easier viewing. Black points
are estimated means from the fitted model, pooling over participants. Black
points and line ranges represent posterior medians and 95% credible intervals,
respectively.

:::

\newpage

::: {#suppfig-tasks-intrinsiceffort suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_task_rank_intrinsiceffort)
```

Model-estimated means for the question "How important is it that effort goes
into this task?" across all 20 tasks in the pilot study. Grey points
represent participant responses to the question, jittered for easier viewing.
Black points are estimated means from the fitted model, pooling over
participants. Black points and line ranges represent posterior medians and 95%
credible intervals, respectively.

:::

\newpage

::: {#suppfig-tasks-extrinsiceffort suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_task_rank_extrinsiceffort)
```

Model-estimated means for the question "How important is it that others see the
effort that goes into this task?" across all 20 tasks in the pilot study.
Grey points represent participant responses to the question, jittered for easier
viewing. Black points are estimated means from the fitted model, pooling over
participants. Black points and line ranges represent posterior medians and 95%
credible intervals, respectively.

:::

\newpage

# Supplementary Tables

::: {.landscape}

::: {#supptbl-tasks supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_tasks) %>%
  flextable() %>%
  theme_apa() %>%
  width(width = c(3, rep(1, 5))) %>%
  line_spacing() %>%
  align(i = NULL, j = 1, align = "left", part = "all")
```

Tasks included in the studies.

:::

:::

::: {.landscape}

::: {#supptbl-essay-answers-social-study3 supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_essay_answers_study3) %>%
  slice(1:3) %>%
  flextable() %>%
  theme_apa() %>%
  width(width = c(1, 8)) %>%
  line_spacing() %>%
  align(i = NULL, j = NULL, align = "left", part = "all")
```

Pre-generated essay answers to the social prompt in Study 3.

:::

:::

::: {.landscape}

::: {#supptbl-essay-answers-nonsocial-study3 supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_essay_answers_study3) %>%
  slice(4:6) %>%
  flextable() %>%
  theme_apa() %>%
  width(width = c(1, 8)) %>%
  line_spacing() %>%
  align(i = NULL, j = NULL, align = "left", part = "all")
```

Pre-generated essay answers to the non-social prompt in Study 3.

:::

:::

::: {.landscape}

::: {#supptbl-essay-comprehension-study3 supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_essay_comprehension_study3) %>%
  flextable() %>%
  theme_apa() %>%
  autofit()
```

Reading times and comprehension rates for the essay answers in Study 3. Expected
reading times were calculated based on an estimated reading speed of 275 words
per minute. Comprehension rates are the percentage of participants who answered
the comprehension question correctly.

:::

:::

::: {#supptbl-manipulation-check-study3 supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_manipulation_check_study3) %>%
  flextable() %>%
  theme_apa() %>%
  autofit()
```

Percentage of participants in Study 3 who passed the manipulation check and
reported that they believed the manipulation, split by condition.

:::

::: {.landscape}

::: {#supptbl-treatment-diffs-person-study3 supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_treatment_diffs_effects_study3) %>%
  dplyr::select(c(1, 7:11)) %>%
  mutate(
    group = c(
      rep("Effect of outsourcing type (Task type = Social)", 3),
      rep("Effect of outsourcing type (Task type = Non-social)", 3),
      rep("Effect of task type", 3),
      rep("Interaction effect", 3)
    )
  ) %>%
  as_grouped_data(groups = "group") %>%
  as_flextable(hide_grouplabel = TRUE) %>%
  theme_apa() %>%
  width(width = c(2, rep(1.5, 5))) %>%
  bold(i = ~ !is.na(group)) %>%
  add_header_row(values = c("", "Response"), colwidths = c(1, 5)) %>%
  align(i = 1, j = NULL, align = "center", part = "header") %>%
  align(i = NULL, j = 1, align = "left", part = "all") %>%
  line_spacing() %>%
  fontsize(size = 8)
```

Pairwise contrasts for character evaluations in Study 3. Numbers reflect
differences in marginal means on a 7-point Likert scale. Estimates are pooled
over essay answers. The bottom rows represent the interactions between
outsourcing type and task type. Main numbers are posterior medians, numbers in
the square brackets are 95% credible intervals.

:::

:::

::: {.landscape}

::: {#supptbl-text-analysis-study6 supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_text_analysis_study6) %>%
  flextable() %>%
  theme_apa() %>%
  autofit()
```

Pairwise comparisons of word frequencies between conditions in Study 6. LL = log
likelihood.

:::

:::

# Supplementary References

Boyd, R. L. (2019). BUTTER: Basic unit-transposable text experimentation
resource. Available from <https://www.butter.tools/>

Gregson, R., Piazza, J., & Boyd, R. L. (2022). 'Against the cult of veganism':
Unpacking the social psychology and ideology of anti-vegans. *Appetite*, *178*,
106143. doi:[10.1016/j.appet.2022.106143](https://doi.org/10.1016/j.appet.2022.106143)

Rayson, P., & Garside, R. (2000). Comparing corpora using frequency profiling.
*Proceedings of the Workshop on Comparing Corpora - Volume 9*, 1–6. Presented in
Hong Kong. doi:[10.3115/1117729.1117730](https://doi.org/10.3115/1117729.1117730)