### 2025-06-23 manuscript.qmd

---
title: "Negative Perceptions of Outsourcing to Artificial Intelligence"
shorttitle: "Outsourcing to AI"
author:
  - name: Scott Claessens
    corresponding: false
    orcid: 0000-0002-3562-6981
    email: scott.claessens@gmail.com
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: UK
        postal-code: CT2 7NP
    role:
      - conceptualization
      - data curation
      - formal analysis
      - investigation
      - methodology
      - visualization
      - writing
      - editing
  - name: Pierce Veitch
    corresponding: false
    orcid: 0009-0005-3364-7470
    email: pv201@kent.ac.uk
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: UK
        postal-code: CT2 7NP
    role:
      - formal analysis
      - methodology
      - editing
  - name: Jim A.C. Everett
    corresponding: true
    orcid: 0000-0003-2801-5426
    email: j.a.c.everett@kent.ac.uk
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: UK
        postal-code: CT2 7NP
    role:
      - conceptualization
      - funding acquisition
      - methodology
      - supervision
      - editing
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    study-registration: This pre-print is currently not yet peer-reviewed, and may differ from the final version.
    data-sharing: Word count = 10,574 words.
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    gratitude: ~
    authorship-agreements: ~
abstract: As artificial intelligence (AI) tools become increasingly integrated into daily life, people are beginning to outsource not only professional tasks but also socio-relational ones. Large language models like ChatGPT can generate wedding vows, speeches, and personal messages, raising questions about how individuals who use AI for such tasks are perceived by others. In this paper, we conduct six pre-registered studies with British participants (N = 3,935) to understand how people view those who expend less effort by outsourcing tasks to AI in different ways. We show that outsourcing makes us think more negatively about not only the person and their motivations, but also the outsourced work itself. Advancing previous work on effort, we show that the specific kind of task that people outsource -- and therefore expend less effort on -- matters; that reduced effort does not lead to uniformly negative character perceptions across different tasks; and that effort is important because it signals second-order judgments that people are being less authentic and care less about the task. Our research highlights the way that AI use shapes our perceptions of people, raising key philosophical questions about efficiency, authenticity, and social ties in a world filled with AI-mediated interactions.
keywords: [artificial intelligence, person perception, outsourcing, effort, trust]
bibliography: bibliography.bib
engine: knitr
format:
  apaquarto-docx:
    fig-dpi: 300
    fig-align: center
execute:
  echo: false
  warning: false
  error: false
crossref:
  custom:
    - kind: float
      key: suppfig
      latex-env: suppfig
      reference-prefix: Supplementary Figure
      space-before-numbering: true
      latex-list-of-description: Supplementary Figure
      caption-location: bottom
    - kind: float
      key: supptbl
      latex-env: supptbl
      reference-prefix: Supplementary Table
      space-before-numbering: true
      latex-list-of-description: Supplementary Table
      caption-location: top
floatsintext: true
---

```{r}
library(flextable)
library(targets)
library(tidyverse)
```

The widespread release of generative AI language models has transformed daily 
life, offering the potential to perform a variety of tasks more efficiently and,
in some cases, with greater effectiveness than by doing them oneself. But as AI
becomes more widely available, people are not only using it to assist them with
things like preparing dinner recipes, writing data analysis code, and planning
daily schedules. Increasingly, AI might be used beyond routine or technical
domains to instead assist in tasks that are more socio-relational in nature,
like writing wedding vows, apology notes, and love letters. Anecdotal evidence
suggests that not only is AI-outsourcing of this kind already happening, but
that it potentially has serious effects on how we judge others. In a recent
Reddit post, a disgruntled newlywed tells the story of her husband using ChatGPT
to write his wedding vows, expressing her discomfort with outsourcing something
to AI that, to her, is deeply meaningful and a reflection of their love for one 
another [@miramar0]. Other reports describe people's negative reactions when
they learn that their romantic partner used ChatGPT to write them an apology 
note [@Tait2024] or even to break up with them [@Anderson2025]. Outsourcing 
tasks -- especially socio-relational ones -- to AI tools may be efficient, but 
could have negative consequences for person perception.

There is nothing new, in principle, about outsourcing tasks. For hundreds of 
years, personal assistants have organized daily schedules, recipe-books have 
provided meal plans, and guidebooks have created travel itineraries. In the 
socio-relational domain, ghostwriters have long-existed, and the internet is 
abound with professional paid services for writing wedding vows and personal 
speeches. AI merely supercharges what is an ancient human impulse: the push to 
reduce mental energy by outsourcing parts of our work onto people, books, tools,
or systems. But even if outsourcing is an old phenomenon, the rapid shift in 
availability and use of AI models has fundamentally changed the ease with which
people can outsource work, what kinds of tasks they can outsource, and the way
in which they can outsource. These new developments in society mean that even as
an old phenomenon in new clothes, there is much we still need to know about
outsourcing.

First, we need to know how people who outsource tasks to AI specifically are 
perceived. We know that people are increasingly using large language models 
(LLMs) for a wide variety of tasks [@UKGov2024]. Due to their ubiquity, perhaps 
outsourcing to LLMs might not lead to negative perceptions? We are sceptical. We 
know that people dislike it when others "free ride" or reduce effort while 
benefiting from collective resources [e.g. @Cubitt2011; @Kerr1983] and that 
people's outputs are perceived as more valuable the more effort was ostensibly 
put into them [@Kruger2004]. Moreover, exertion of effort is deemed morally 
admirable and is rewarded, even in situations where effort does not directly 
generate additional product, quality, or economic value, suggesting that effort 
itself is moralized [@Celniker2023]. While outsourcing and reduced effort are
not synonymous, it seems the core social psychological processes are likely to
be similar: when someone outsources a task instead of doing it themselves, they
are expending less effort to achieve their goal, and people value effort. 
Indeed, some work shows that describing someone as using AI for a relational 
task led to the perception they expended less effort and were less satisfied 
with their relationship [@Liu2024]. Other work finds that people receive more 
negative social evaluations if they use AI to complete assignments at school and
at work [@Reif2025; @Roth2025]. We do not know, however, how outsourcing -- and 
therefore, presumably, expending less effort -- might be different for a human 
or an AI, and how these processes will occur in the myriad ways in which people 
can and do use AI to help them complete tasks.

Second, we need to know whether the *type* of task that people are outsourcing 
matters. One might expect outsourcing to be perceived negatively regardless of 
the type of task being outsourced -- if effort is generally moralized, then the 
domain in which it is expended (or not) should have little impact. However, 
there are reasons to expect differences between social tasks like writing vows 
and non-social tasks like writing computer code. We know that different norms,
standards, and expectations can be applied to social and non-social tasks and 
exchanges [e.g. @Fiske1992; @Heider1958; @Malle2022]. Moreover, from a 
philosophical perspective, it often matters not only *whether* something is 
done, but *how* it is done [@Aristotle2009; @Hursthouse2023; @Stohr2006]. An 
apology is not just about hearing someone say "I am sorry", but seeing genuine 
regret; a love letter is not just about hearing someone say "I love you", but 
seeing depth of emotion; and a bereavement letter is not just about hearing 
someone say "I am sorry for your loss", but seeing an understanding for the 
powerful human experience of loss. There is, perhaps especially for social 
tasks, value not only in the outcome of doing something, but the *process* too 
[@Goodman2010]. Previous work on the role of effort, however, has tended to 
focus on a limited number of tasks, looking at how people judge the quality of 
products like paintings, poems, or suits of armour based on how much effort is 
perceived [@Kruger2004] or how people judge others who expend more or less 
effort in employment and physical effort like marathon running [@Celniker2023]. 
Yet the number of tasks that people can outsource –- and therefore, presumably, 
expend less effort –- is vast, and these tasks can vary in different ways. 
Different kinds of tasks that people can outsource can be more or less socially 
relevant, more or less consequential, and more or less likely to impact others: 
it is likely that whether someone puts effort into a socially-relevant task like 
planning a friend's birthday party or writing an apology letter would have 
stronger effects on perceived morality than whether they put effort into writing 
computer code or planning a daily schedule – yet this reminds untested. To 
understand any potential negative effects of outsourcing -- to AI or to humans 
-- we must therefore look at a broad range of non-social and social tasks, 
rather than draw broad conclusions based on a few use cases.

Third, we need to know how outsourcing may shape different kinds of social 
perceptions. People can judge others on separate dimensions of warmth and 
competence [e.g., @Abele2021; @Fiske2007] as well as on dimensions of morality 
and trustworthiness [@Goodwin2014]. It remains unclear how outsourcing to AI 
might lead to differential character judgments across these different 
dimensions. For example, @Kruger2004 focus on judgments of the output, but not 
the person, and while @Celniker2023 include the core social dimensions of 
warmth, morality, and trustworthiness, we do not know how these character 
ratings would be affected by outsourcing to an AI (as well as a human) and how 
this might differ from laziness as a distinct trait. And, most importantly, we 
do not know how outsourcing different kinds of tasks might influence different 
perceptions: does outsourcing in the social domain lead to particularly negative 
effects on warmth, morality, and trustworthiness?

Fourth, we need to know how different *ways* of outsourcing to AI influence 
negative perceptions. Someone who "fully" outsources a task to AI by simply 
giving it a prompt and copying the output word-for-word might be perceived very
differently to someone who gives the AI a prompt, revises the work accordingly,
and finishes it themselves -- using AI as a *collaborative tool*, rather than as
a replacement. Similarly, someone could deceive others about their use of AI or
be perfectly honest about it, and someone could use AI because they are in a 
rush or because the task is important to them and they want to get it right. 
While it seems reasonable to assume that "fully" outsourcing would be perceived 
worse than using AI as a collaborative tool, and that not acknowledging AI use 
would be perceived worse than being honest about it, it remains unclear how much 
this reduces negative perceptions: if someone uses AI in the "best" way (by 
using it as a collaborative tool, being open about their use of AI, and having a 
good reason for using AI), would they still suffer negative social consequences 
from doing so?

Fifth, we need to understand *why* outsourcing to AI, and therefore expending 
less effort, might have these effects. Previous work has focused on how 
expending less effort directly leads to negative perceptions of others 
[@Celniker2023]. But this raises the question of *why* effort is seen as 
important and what exactly it is signalling to others, beyond one's general 
cooperative intent. It is possible that outsourcing leads to negative 
perceptions because the lack of effort spent on the task signals something more 
fundamental about how authentic one is and how much one cares about the task: 
when someone chooses to outsource a love letter to an AI, they might be seen as 
valuing that love letter and what it represents less. It could be this 
second-step order of perceptions that is the key driver of negative perceptions, 
and this, in turn, could especially be so for the kinds of tasks that are more 
socially relevant.

## Present Research

In this paper, we build on classic social psychological work on character 
inferences from reduced effort to understand how people view others who
outsource different kinds of tasks, in different ways, for different reasons, to
AI. Across six pre-registered experiments with British participants, we seek
not only to understand how reduced effort through AI-outsourcing might shape
perceptions of others, but also advance previous theoretical work on the
moralization of effort [@Celniker2023]. Moving beyond a focus on how effort 
itself shapes character judgments, we seek to provide deeper theoretical nuance 
into why effort is important: not only because effort itself is intrinsically 
moralized, but that expending effort gives rise to second-order inferences that 
independently influence character judgments.

In Study 1, we explore the perceptions of people who outsource a range of tasks 
to AI or another person, compared to when they complete the tasks by themselves.
In Study 2, we then look at the effects of task type, AI use, and honesty. We 
explore how people perceive others who outsource different kinds of tasks with 
different levels of social relevance (e.g., from daily schedules, computer code 
and dinner recipes to wedding vows, apology letters, or bereavement cards), 
manipulating whether people use AI as a collaborative tool or "fully" outsource 
to AI and whether they are honest or deceptive about their use of AI. After 
turning to look at perceptions of both outsourcers and the outsourced work in 
Study 3, in Studies 4-6 we probe why outsourcing may have negative effects on 
how we evaluate others. In Study 4, we test potential mechanisms of perceived 
effort and authenticity by looking at how people evaluate others who either 
spend a lot or little time crafting the AI prompts, and who either outsource to 
a generic or personalized AI. In Study 5, we test the potential mechanism of 
perceived importance in the task by manipulating people's reasons for using AI 
-- either because they wanted to save time or because they cared about the task 
and thought that AI would improve their work. Finally, in Study 6, we bring 
these different potential mechanisms together to explore the different pathways 
that influence the relationship between outsourcing and negative perceptions,  
focusing on perceived effort, authenticity, and care in the task.

# Study 1

## Methods

### Ethical Approval

Ethical approval was granted for all studies in this paper by the REDACTED 
Psychology Research Ethics Panel. Participants in all studies provided
informed consent and were debriefed after the study.

### Participants

We conducted a power simulation to determine our target sample size. The
simulation suggested that a sample size of 150 participants per condition
(overall *n* = 450 for three conditions) would be required to detect a small
difference between conditions (Cohen's *d* ≈ 0.20) with above 80% power.

```{r}
study1_sample <-
  tar_read(study1_data) %>%
  group_by(id) %>%
  summarise(
    gender = unique(gender),
    age = unique(age),
    chatgpt_used = unique(chatgpt_used)
    )
```

We recruited a convenience sample of 500 participants from the United Kingdom
through the online platform Prolific (<https://www.prolific.com/>). After 
excluding participants who failed our pre-treatment attention check, we were 
left with a final sample of `r nrow(study1_sample)` participants 
(`r sum(study1_sample$gender == "Female")` female; 
`r sum(study1_sample$gender == "Male")` male;
`r sum(study1_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(study1_sample$gender, "Prefer") | is.na(study1_sample$gender))`
undisclosed gender; mean age = `r round(mean(study1_sample$age), 2)` years).
`r round(mean(study1_sample$chatgpt_used == "Yes", na.rm = TRUE) * 100, 0)`% of
these participants reported having used ChatGPT before.

### Design

We randomly allocated participants into one of three conditions in a
between-subjects design: (*i*) the control condition, (*ii*) the AI outsourcing
condition, or (*iii*) the human outsourcing condition. These conditions
determined how scenarios were presented to participants.

### Procedure

We presented participants with six scenarios. Each scenario described a person
completing a task, such as writing computer code or writing a love letter. The
six tasks were randomly drawn from a larger set of 20 tasks (see @supptbl-tasks
for the full list of tasks). For each scenario, we told participants:

- *Control condition*: "In order to complete this task, [the person] works on it
by themselves from start to finish."
- *AI outsourcing condition*: "In order to complete this task, [the person] gets
the AI tool ChatGPT to do it for them."
- *Human outsourcing condition*: "In order to complete this task, [the person]
gets someone else to do it for them."

<br>

We then asked participants how well each of the following words described the
person in the scenario: competent, warm, moral, lazy, and trustworthy.
Participants answered these questions on 7-point Likert scales, ranging from
"does not describe [the person] well" to "describes [the person] extremely well".

After the six scenarios, we asked participants several questions about the AI
tool ChatGPT, including their familiarity with ChatGPT, whether they had used
ChatGPT before, how frequently they used ChatGPT, and how trustworthy they
thought ChatGPT was.

### Pre-registration

We pre-registered the study on the Open Science Framework
(<https://osf.io/xhmzk/?view_only=a4da193574d7410ba4d2aa3945a28b05>).

### Statistical Analysis

We fitted Bayesian multivariate multilevel cumulative-link ordinal models to the
data using the *brms* R package. We modelled each character evaluation --
competence, warmth, morality, laziness, and trustworthiness -- as a separate
response variable and included fixed effects for conditions, varying intercepts
for participants, and varying intercepts and slopes for tasks. We used
regularizing priors for all parameters to impose conservatism on parameter
estimates. All models converged normally ($\hat{R}$ ≤ 1.01).

### Transparency and Openness

For all studies in this paper, we report how we determined our sample size, all
data exclusions, all manipulations, and all measures in the studies. All studies
were pre-registered. Analyses for all studies were conducted in R v4.4.2 
[@RCoreTeam]. Visualizations were produced using the *ggplot2* and *patchwork* 
packages [@Wickham2016; @Pedersen2025]. The manuscript was reproducibly 
generated using the *targets* package [@Landau2021] and *quarto* [@Allaire2024].
All survey materials, data, and code to reproduce the analyses and figures in 
this paper can be found here: 
<https://osf.io/xhmzk/?view_only=a4da193574d7410ba4d2aa3945a28b05>

## Results

We found that people who outsourced tasks to AI or other humans were perceived
more negatively than people who completed the tasks themselves
(@fig-treatments-study1). In particular, people who outsourced were perceived as 
lazier and less competent, with smaller yet detectable differences for 
perceptions of warmth, morality, and trustworthiness
(@tbl-treatment-diffs-study1). Across all measures, outsourcing to other humans 
was perceived more negatively than outsourcing to AI.

```{r, fig.height=4, fig.width=6}
#| label: fig-treatments-study1
#| fig-cap: Overall Character Evaluations in Study 1
#| fig-align: center
#| apa-note: Participants in the control condition, the AI outsourcing condition, and the human outsourcing condition evaluated people in the scenarios on (a) competence, (b) warmth, (c) morality, (d) laziness, and (e) trustworthiness. Colored points represent participant responses to the questions, jittered for easier viewing. Black points are estimated marginal means from the fitted model, pooling over participants and tasks. Black points and line ranges represent posterior medians and 95% credible intervals, respectively.

tar_read(plot_treatments_study1)
```

::: {.landscape}

```{r}
#| label: tbl-treatment-diffs-study1
#| tbl-cap: Overall Pairwise Contrasts in Study 1 \vspace{20pt}
#| apa-note: Numbers reflect differences in marginal means on a 7-point Likert scale, pooling over participants and tasks. Main numbers are posterior medians, numbers in the square brackets are 95% credible intervals.

tar_read(table_treatment_diffs_study1) %>%
  flextable() %>%
  theme_apa() %>%
  width(width = c(1.5, rep(1.5, 5))) %>%
  add_header_row(values = c("", "Response"), colwidths = c(1, 5)) %>%
  align(i = 1, j = NULL, align = "center", part = "header") %>%
  align(i = NULL, j = 1, align = "left", part = "all") %>%
  line_spacing() %>%
  fontsize(size = 10)
```

:::

We found that the effects of outsourcing varied across the different tasks,
especially for perceptions of warmth and morality 
(@fig-treatments-tasks-study1). For example, people were perceived as
less warm if they outsourced writing a love letter, but not if they outsourced
writing computer code. Similarly, people were perceived as less moral if they
outsourced writing an apology letter to a friend, but not if they outsourced
writing a dinner recipe. By contrast, the effects of outsourcing on competence,
laziness, and trustworthiness were more consistent across tasks.

::: {.landscape}

```{r, fig.height=5, fig.width=8.5}
#| label: fig-treatments-tasks-study1
#| fig-cap: Variation in the Effects of Outsourcing across Tasks in Study 1 \vspace{-30pt}
#| fig-align: center
#| apa-note: Tasks are ordered from most social (top) to least social (bottom) according to ratings from the pilot study. Point ranges are differences in marginal means on a 7-point Likert scale for the AI outsourcing condition (red) and the human outsourcing condition (blue) compared to the control condition. Points and ranges represent posterior medians and 95% credible intervals, respectively.

tar_read(plot_treatments_tasks_study1)
```

:::

To determine the factors that predict variation across tasks, we incorporated
ratings of tasks from a pilot study (see Supplementary Materials). Participants 
were asked to rate the 20 tasks on several features: whether the task is social, 
requires social skills, impacts others, has important consequences, and requires 
effort. All of these features predicted stronger causal effects of outsourcing 
compared to control (Supplementary Figures [-@suppfig-interactions-study1] 
and [-@suppfig-interaction-pars-study1]). In other words, outsourcing to AI 
or other humans was perceived more negatively for tasks that were rated as being 
socially relevant and impactful.

## Discussion

In Study 1, we conducted our first exploration of how outsourcing to either a 
human or an AI shaped character perceptions across a variety of tasks. Our 
results show that people who outsourced the task instead of completing it 
themselves were seen as less moral, warm, competent, and trustworthy, and more 
lazy. We found that outsourcing a task led to negative character judgments 
regardless of whether the task was outsourced to a human or a machine, though 
predictably we found that outsourcing to a human was seen as worse given that 
LLMs are at least purposefully designed to be used. And finally, we found that 
the kind of task matters: these negative perceptions were not uniform, but were 
particularly found for tasks that were more social and impactful.

In Study 2, we turned to focus specifically on the different ways that people 
might expend less effort by outsourcing to an AI, focusing on the role of type 
of use (using AI as a tool but still finishing the task oneself vs. full 
delegating) and honesty about the use of AI. Would these negative perceptions be 
seen even in the "best case" scenario where someone uses AI as a tool and is 
honest about it?

# Study 2

## Methods

### Participants

We conducted a power simulation to determine our target sample size. The 
simulation suggested that a sample size of 150 participants per condition
(overall *n* = 750 for five conditions) would be required to detect a small 
difference between conditions (Cohen's *d* ≈ 0.20) with above 80% power.

```{r}
study2_sample <-
  tar_read(study2_data) %>%
  group_by(id) %>%
  summarise(
    gender = unique(gender),
    age = unique(age),
    chatgpt_used = unique(chatgpt_used)
    )
```

We recruited a convenience sample of 800 participants from the United Kingdom 
through Prolific. After excluding participants who failed our pre-treatment 
attention check, we were left with a final sample of `r nrow(study2_sample)` 
participants (`r sum(study2_sample$gender == "Female")` female; 
`r sum(study2_sample$gender == "Male")` male;
`r sum(study2_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(study2_sample$gender, "Prefer") | is.na(study2_sample$gender))`
undisclosed gender; mean age = `r round(mean(study2_sample$age), 2)` years).
`r round(mean(study2_sample$chatgpt_used == "Yes", na.rm = TRUE) * 100, 0)`% of
these participants reported having used ChatGPT before.

### Design

We used a "control plus 2x2" between-subjects design. Participants were randomly 
allocated to either the control condition, in which people in the scenarios 
complete the tasks themselves, or one of four experimental conditions, in which 
people in the scenarios use AI to complete the tasks. In the experimental 
conditions, we manipulated whether people in the scenarios used AI as a 
collaborative tool or "fully" outsourced to AI, and whether people were honest 
or deceptive about their use of AI. This resulted in five conditions overall: 
(*i*) the control condition, (*ii*) the tool-honest condition, (*iii*) the 
tool-deception condition, (*iv*) the full-honest condition, and (*v*) the 
full-deception condition.

### Procedure

We presented participants with six scenarios. Each scenario described a person 
completing a task, such as writing computer code or writing a love letter. The 
six tasks were randomly drawn from a larger set of 16 tasks (see @supptbl-tasks 
for the full list of tasks). For each scenario, we first told participants:

- *Control condition*: "In order to complete this task, [the person] works on it 
by themselves from start to finish."
- *Tool outsourcing conditions*: "In order to complete this task, [the person] 
uses the AI tool ChatGPT. They ask ChatGPT to provide ideas, inspiration, and 
feedback, but they edit and rewrite the suggestions and finish the task
themselves."
- *Full outsourcing conditions*: "In order to complete this task, [the person]
uses the AI tool ChatGPT. They copy ChatGPT's output word-for-word, rather than
doing it themselves."

<br>

We then told participants in the experimental conditions:

- *Honest conditions*: "After completing the task, [the person] is asked how
they came up with their ideas. [The person] acknowledges that they used ChatGPT
as a tool / got ChatGPT to do the task for them."
- *Deception conditions*: "After completing the task, [the person] is asked how
they came up with their ideas. [The person] does not acknowledge that they used 
ChatGPT as a tool / got ChatGPT to do the task for them."

<br>

We then asked participants how well each of the following words described the 
person in the scenario: competent, warm, moral, lazy, and trustworthy. 
Participants answered these questions on 7-point Likert scales, ranging from 
"does not describe [the person] well" to "describes [the person] extremely 
well".

After the six scenarios, we asked participants several questions about the AI 
tool ChatGPT, including their familiarity with ChatGPT, whether they had used 
ChatGPT before, how frequently they used ChatGPT, and how trustworthy they 
thought ChatGPT was.

### Pre-registration

We pre-registered the study on the Open Science Framework
(<https://osf.io/xhmzk/?view_only=a4da193574d7410ba4d2aa3945a28b05>).

### Statistical Analysis

We fitted Bayesian multivariate multilevel cumulative-link ordinal models to the
data using the *brms* R package [@Burkner2017]. We modelled each character 
evaluation -- competence, warmth, morality, laziness, and trustworthiness -- as 
a separate response variable and included fixed effects for conditions, varying 
intercepts for participants, and varying intercepts and slopes for tasks. We 
used regularizing priors for all parameters to impose conservatism on parameter 
estimates. All models converged normally ($\hat{R}$ ≤ 1.01).

## Results

We first looked at the overall results averaging over tasks. Across all five 
character evaluations, we found that fully outsourcing to AI (i.e., copying the 
AI output verbatim) was perceived more negatively than using AI as a 
collaborative tool (@fig-treatments-study2; @tbl-treatment-diffs-study2). By 
contrast, we found that deception about AI usage had specific negative effects 
on perceptions of morality and trustworthiness: people who did not acknowledge 
their use of AI were perceived as less moral and less trustworthy. We did not 
find any interaction effects between full outsourcing and deception.

```{r, fig.height=6, fig.width=6}
#| label: fig-treatments-study2
#| fig-cap: Overall Character Evaluations in Study 2
#| fig-align: center
#| apa-note: Participants in the control condition and four AI outsourcing conditions evaluated people in the scenarios on (a) competence, (b) warmth, (c) morality, (d) laziness, and (e) trustworthiness. Colored points represent participant responses to the questions, jittered for easier viewing. Black points are estimated marginal means from the fitted model, pooling over participants and tasks. Black points and line ranges represent posterior medians and 95% credible intervals, respectively.

tar_read(plot_treatments_study2)
```

::: {.landscape}

```{r}
#| label: tbl-treatment-diffs-study2
#| tbl-cap: Overall Pairwise Contrasts in Study 2 \vspace{20pt}
#| apa-note: Numbers reflect differences in marginal means on a 7-point Likert scale, pooling over participants and tasks. The bottom row represents the interaction between full outsourcing and deception (i.e., the difference between the differences in the rows above). Main numbers are posterior medians, numbers in the square brackets are 95% credible intervals.

tar_read(table_treatment_diffs_effects_study2) %>%
  mutate(
    group = c(
      rep("Comparison to control", 4),
      rep("Effect of full outsourcing", 2),
      rep("Effect of deception", 2),
      "Interaction effect"
    )
  ) %>%
  as_grouped_data(groups = "group") %>%
  as_flextable(hide_grouplabel = TRUE) %>%
  theme_apa() %>%
  width(width = c(2, rep(1.5, 5))) %>%
  bold(i = ~ !is.na(group)) %>%
  add_header_row(values = c("", "Response"), colwidths = c(1, 5)) %>%
  align(i = 1, j = NULL, align = "center", part = "header") %>%
  align(i = NULL, j = 1, align = "left", part = "all") %>%
  line_spacing() %>%
  fontsize(size = 10)
```

:::

The effects of outsourcing to AI varied across the different tasks, especially 
for perceptions of warmth and morality (@fig-treatments-tasks-study1). For 
example, people who used AI for social tasks, such as writing an apology letter 
or a bereavement card, were perceived as less warm, less moral, and lazier 
compared to people who completed the task themselves. This was true even if the 
person used AI as a tool and was honest about their usage of AI. By contrast, we 
observed weaker effects of outsourcing for non-social tasks like writing 
computer code or planning a syllabus.

::: {.landscape}

```{r, fig.height=5, fig.width=8.5}
#| label: fig-treatments-tasks-study2
#| fig-cap: Variation in the Effects of Outsourcing across Tasks in Study 2 \vspace{-30pt}
#| fig-align: center
#| apa-note: Tasks are ordered from most social (top) to least social (bottom) according to ratings from a pilot study. Point ranges are differences in marginal means on a 7-point Likert scale for the honest conditions (red) and deception conditions (blue) compared to the control condition. Upper panels refer to the tool outsourcing conditions, and lower panels refer to the full outsourcing conditions. Points and ranges represent posterior medians and 95% credible intervals, respectively.

tar_read(plot_treatments_tasks_study2)
```

:::

To determine the factors that predict variation across tasks, we incorporated 
ratings of tasks from Study 1. We found stronger causal effects of outsourcing 
to AI compared to the control for tasks that were rated as more social, 
requiring more social skills, impacting others more, having more important 
consequences, and requiring effort (Supplementary Figures 
[-@suppfig-interactions-study2] and [-@suppfig-interaction-pars-study2]).

## Discussion

In Study 2, we looked at how people who outsourced to AI in different ways were 
perceived across a broad range of social and non-social tasks. In line with our 
predictions, we found that "fully" outsourcing to AI was perceived more 
negatively than using AI as a collaborative tool, particularly for 
socio-relational tasks. We also found, predictably, that people were seen as 
less moral and less trustworthy if they did not acknowledge their use of AI. 
Importantly, though, we show that even using AI in the "best" way -- only as a 
tool and being honest about one's usage -- still led to negative social 
perceptions for the more socio-relational tasks like writing a love letter, an 
apology, or wedding vows.

In Study 3, we investigate whether these negative perceptions extend to the work 
itself and remain after seeing the output. It could be, for example, that 
someone is perceived badly for using ChatGPT to write their bereavement card, 
but the writing itself is seen as equally well-written and authentic, if not 
more so, than if the person had written the card themselves. Indeed, evidence 
suggests that text generated by ChatGPT is rated as higher quality than 
human-written text [@Noy2023]. Moreover, it is possible that seeing appropriate 
output could mitigate negative perceptions by highlighting how the AI can in 
fact perform the task well. We explored these possibilities in Study 3.

# Study 3

## Methods

### Participants

We conducted a power simulation to determine our target sample size. The 
simulation suggested that a sample size of 125 participants per condition 
(overall *n* = 750 for six conditions) would be required to detect a 
small-to-medium difference between conditions (Cohen's *d* ≈ 0.40) with above 
80% power.

```{r}
study3_sample <-
  tar_read(study3_data) %>%
  group_by(id) %>%
  summarise(
    gender = unique(gender),
    age = unique(age),
    chatgpt_used = unique(chatgpt_used)
    )
```

We recruited a convenience sample of 800 participants from the United Kingdom 
through Prolific. After excluding participants who failed our pre-treatment 
attention check, we were left with a final sample of `r nrow(study3_sample)` 
participants (`r sum(study3_sample$gender == "Female")` female; 
`r sum(study3_sample$gender == "Male")` male;
`r sum(study3_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(study3_sample$gender, "Prefer") | is.na(study3_sample$gender))`
undisclosed gender; mean age = `r round(mean(study3_sample$age), 2)` years).
`r round(mean(study3_sample$chatgpt_used == "Yes", na.rm = TRUE) * 100, 0)`% of
these participants reported having used ChatGPT before.

### Design

We randomly allocated participants to one of six conditions in a 3x2 
between-subjects design. We manipulated the type of outsourcing: (*i*) no 
outsourcing control, (*ii*) using AI as a tool, and (*iii*) fully outsourcing to
AI. Here, in contrast to Studies 1 and 2, we also explicitly manipulated whether 
the task prompt was social or non-social.

### Procedure

We told participants that they would read and evaluate a short piece of writing 
from "another participant". In reality, we generated the writing using ChatGPT 
version 4o. We asked ChatGPT to generate a 300 word response to the prompt and 
to write convincingly like a real human. We then edited the text to appear more 
human-like by, for example, removing classic AI markers like dashes and 
concluding sentences and ensuring that the information was not too generic, such 
that the writing could reasonably be attributed to both a human and AI.

The prompt for the piece of writing varied between conditions:

- *Social conditions*: "Please write a description of a close family member or 
friend, explaining why they are special to you."
- *Non-social conditions*: "Please write a short description of a book, TV show,
or film of your choice."

<br>

We explained that the "other participant" was asked several questions about how
they produced their answer, including whether or not they used an AI tool like
ChatGPT. We explained that the participant was encouraged to be honest and told
that they would be paid regardless. The response from the "other participant" 
varied between conditions:

- *Control conditions*: "The participant reported that they did not use any AI 
tool like ChatGPT. Instead, they worked on the response themselves from start to
finish."
- *Tool outsourcing conditions*: "The participant reported using ChatGPT to 
provide ideas, inspiration, and feedback. The participant told us that they 
edited and rewrote ChatGPT's suggestions and finished writing the response 
themselves."
- *Full outsourcing conditions*: "The participant reported using ChatGPT to 
complete the task. The participant told us that they copied ChatGPT's output
word-for-word, rather than producing the response themselves."

<br>

Next, we presented participants with a randomly-chosen pre-generated essay 
answer to the prompt (see Supplementary Tables 
[-@supptbl-essay-answers-social-study3] and 
[-@supptbl-essay-answers-nonsocial-study3] for full essay answers). In the 
social conditions, the answer either referred to the participant's father, their
sister, or their best friend. In the non-social conditions, the answer either
referred to the book The Hobbit, the TV show Buffy the Vampire Slayer, or the
film Titanic. Reading times and responses to a follow-up comprehension question
suggested that participants read the essay answers in sufficient detail (see 
@supptbl-essay-comprehension-study3).

Finally, we asked participants about their perceptions of the essay answer and 
the "other participant". We asked how well-written, meaningful, and authentic 
they thought the answer was (7-point Likert scales), what letter grade they 
would give the answer (A-E), and how much they would hypothetically reward the 
other participant for their work (from £0.00 to £1.00). We also asked how well 
each of the following words described the other participant: competent, warm, 
moral, lazy, and trustworthy (7-point Likert scales).

At the end of the study, we gave participants a manipulation check and asked
them whether they believed the manipulation. Almost all participants correctly 
reported the condition that they were in and most participants stated that they 
believed the essay response was written in the way we described, suggesting that 
the manipulation was successful (see @supptbl-manipulation-check-study3). We 
also asked participants several questions about ChatGPT.

### Pre-registration

We pre-registered the study on the Open Science Framework[^1].

[^1]: Due to a technical error with archiving this pre-registration on the Open
Science Framework, the timestamp for the registration was lost. However, on our
OSF project (<https://osf.io/xhmzk/?view_only=a4da193574d7410ba4d2aa3945a28b05>), it is possible to view our 
pre-registration document file and its timestamped upload date.

### Statistical Analysis

We fitted two Bayesian multilevel models to the data. The first model was a 
multivariate cumulative-link ordinal model including all Likert scales as 
separate response variables. The second model was a zero-one-inflated-beta model
applied specifically to the reward question, which was a slider scale varying
between 0 and 1. For both models, we included fixed effects for the interaction
between outsourcing type and task type and varying intercepts and slopes for
essay answers. We used regularizing priors for all parameters to impose 
conservatism on parameter estimates. All models converged normally 
($\hat{R}$ ≤ 1.01).

## Results

We first looked at character evaluations. We found that even when provided with 
concrete output, people were still perceived more negatively across all 
character evaluations if they outsourced the writing task to AI, either by using 
ChatGPT as a collaborative tool or by copying the text from ChatGPT verbatim 
(@suppfig-treatments-person-study3; @supptbl-treatment-diffs-person-study3). 
In contrast to Study 2, however, we did not find any differences in character 
evaluations between the tool outsourcing and full outsourcing conditions. We did 
not find differences in character evaluations between social and non-social 
tasks and did not find any interaction effects.

Turning to evaluations of the work itself, we found that the AI-outsourced work 
(either outsourced by using AI as a collaborative tool or fully outsourced) was 
judged as being equally well written to the work in the control condition 
(@fig-treatments-work-study3; @tbl-treatment-diffs-work-study3). This is in line 
with the writing indeed being identical in all conditions. Interestingly, 
however, we found that essay responses that were ostensibly generated using AI 
were perceived as less meaningful and less authentic compared to essay responses 
ostensibly written by a human. Participants also marked AI-generated essays with 
a lower grade and rewarded AI-generated essays with a lower hypothetical 
monetary bonus. In contrast to Study 2, we did not find differences in 
perceptions of the work between the tool outsourcing and full outsourcing 
conditions, except for the reward question, where fully outsourced essays (i.e., 
essays copied verbatim from ChatGPT) were rewarded £0.23 less than essays 
generated using AI as a collaborative tool. We did not find any differences 
between social and non-social tasks and did not find any interaction effects.

```{r, fig.height=5, fig.width=7}
#| label: fig-treatments-work-study3
#| fig-cap: Perceptions of the Work in Study 3
#| fig-align: center
#| apa-note: Participants in the control condition, the tool outsourcing condition, and the full outsourcing condition evaluated the essay response to the writing task. Participants rated whether the essay response was (a) well-written, (b) meaningful, and (c) authentic. Participants also (d) graded the work and (e) rewarded the work with a hypothetical monetary bonus. Jittered points represent participant responses to the questions, split by whether the writing task was a non-social task (red) or a social task (blue). Point ranges are estimated marginal means from the fitted model, pooling over essay answers. Points and line ranges represent posterior medians and 95% credible intervals, respectively.

tar_read(plot_treatments_work_study3)
```

::: {.landscape}

```{r}
#| label: tbl-treatment-diffs-work-study3
#| tbl-cap: Pairwise Contrasts for Perceptions of the Work in Study 3 \vspace{20pt}
#| apa-note: Numbers reflect differences in marginal means on either a 7-point Likert scale (well-written, meaningful, authentic), a 5-point ordinal grade scale (grade), or a 0-1 sliding scale (reward). Estimates are pooled over essay answers. The bottom rows represent the interactions between outsourcing type and task type. Main numbers are posterior medians, numbers in the square brackets are 95% credible intervals.

tar_read(table_treatment_diffs_effects_study3) %>%
  dplyr::select(1:6) %>%
  mutate(
    group = c(
      rep("Effect of outsourcing type (Task type = Social)", 3),
      rep("Effect of outsourcing type (Task type = Non-social)", 3),
      rep("Effect of task type", 3),
      rep("Interaction effect", 3)
    )
  ) %>%
  as_grouped_data(groups = "group") %>%
  as_flextable(hide_grouplabel = TRUE) %>%
  theme_apa() %>%
  width(width = c(2.5, rep(1.3, 5))) %>%
  bold(i = ~ !is.na(group)) %>%
  add_header_row(values = c("", "Response"), colwidths = c(1, 5)) %>%
  align(i = 1, j = NULL, align = "center", part = "header") %>%
  align(i = NULL, j = 1, align = "left", part = "all") %>%
  line_spacing() %>%
  fontsize(size = 8)
```

:::

## Discussion

In Study 3, we turned to look at how people perceived both the outsourcer and 
the outsourced work when given specific output in a social or non-social task 
that was described as being produced independently by a person, produced by a 
person in collaboration with AI as a tool, or outsourced in full to AI. We find 
that our results generalize from character judgments to perceptions of the work 
itself: text purportedly generated using AI was perceived to be less meaningful, 
less authentic, and less reward-worthy compared to the same text described as 
human-generated.

Surprisingly, we found no differences in the effect of AI-outsourcing between 
social and non-social tasks. This may be due to the particular tasks we chose. 
Writing *about* someone close to you is not quite the same as writing something 
*for* someone close to you, as is the case with wedding vows, love letters, and 
bereavement cards. We also found no differences between the tool and full 
outsourcing conditions, aside from the lower rewards given to participants in 
the latter condition. It is possible that because the set-up described to 
participants was of another participant who was asked to produce work on 
Prolific and then admitted they used AI, participants saw any kind of AI use as 
violating an implicit contract between the survey requester and respondent and 
judged them negatively accordingly.

In Study 4, we turn to explore potential mechanisms driving our effects. We 
assume that effort may play a role, since perceived effort is often used as a 
signal of one's moral character [@Cubitt2011] and cooperative intent 
[@Celniker2023]. Study 3 also suggested a role of authenticity: in line with 
work on the psychological importance of authenticity [@Newman2019], people who 
outsource to AI may be perceived as producing work that is less authentically 
their own, leading to negative evaluations. To what extent would negative 
perceptions of outsourcing persist when someone is described as putting a lot of
effort into the prompts that they fed to the LLM, and when someone is described
as using a personalized -- and therefore more "authentic" -- LLM rather than a 
generic LLM like ChatGPT? To explore this, in Study 4 we experimentally 
manipulate (1) how much effort someone puts into the task and (2) whether they 
outsource the task to a standard LLM like ChatGPT or a personalized LLM trained 
specifically on their own prior writings (and so therefore producing work that 
is more authentically "theirs"). We expected negative perceptions of outsourcing
to be mitigated when the person uses a personalized LLM and expends significant
effort on formulating prompts for the AI.

# Study 4

## Methods

### Participants

```{r}
study4_sample <-
  tar_read(study4_data) %>%
  group_by(id) %>%
  summarise(
    gender = unique(gender),
    age = unique(age),
    chatgpt_used = unique(chatgpt_used),
    effort = unique(effort),
    authentic = unique(authentic)
    )
```

We used the same power estimate from Study 2 to determine our target sample size
of *n* = 750 (150 participants in each of five conditions). We recruited a 
convenience sample of 802 participants from the United Kingdom through Prolific.
After excluding participants who failed our pre-treatment attention check, we
were left with a final sample of `r nrow(study4_sample)` 
participants (`r sum(study4_sample$gender == "Female")` female; 
`r sum(study4_sample$gender == "Male")` male;
`r sum(study4_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(study4_sample$gender, "Prefer") | is.na(study4_sample$gender))`
undisclosed gender; mean age = `r round(mean(study4_sample$age), 2)` years).
`r round(mean(study4_sample$chatgpt_used == "Yes", na.rm = TRUE) * 100, 0)`% of
these participants reported having used ChatGPT before.

### Design

We used the same "control plus 2x2" between-subjects design as in Study 2. In 
the experimental conditions, we manipulated whether people in the scenarios used
a standard or personalized AI model, and whether people put more or less effort
into the task. This resulted in five conditions overall: (*i*) the control 
condition, (*ii*) the standard-low-effort condition, (*iii*) the 
standard-high-effort condition, (*iv*) the personalized-low-effort condition, 
and (*v*) the personalized-high-effort condition. Our authenticity manipulation 
was inspired by recent psychological work looking at the credit-blame asymmetry 
in AI use [@Earp2024], showing that people receive more personal credit for 
their work when they use an AI model trained on their own prior writings.

### Procedure

The procedure was mostly identical to Study 2 to allow us to explore effects 
across a range of tasks, but we updated the study preamble and the presentation 
of the scenarios. For participants in the personalized AI conditions, we 
expanded the study preamble to explain that personalized AI models were trained
on people's own prior writings and "tailored to each specific person and their 
own thoughts, feelings, and values". Then in the scenarios, we told participants
in the experimental conditions:

- *Standard AI conditions*: "In order to complete this task, [the person] uses 
the AI tool ChatGPT."
- *Personalized AI conditions*: "In order to complete this task, [the person]
uses a personalized AI tool."

<br>

We then told participants:

- *Low effort conditions*: "[The person] quickly gives the AI a rushed prompt 
and uses its first output."
- *High effort conditions*: "[The person] carefully gives the AI several 
detailed prompts and, after multiple rounds of changes, uses its resulting 
output."

<br>

At the end of the study, we asked participants to choose which of these was more
authentic and effortful, respectively. 
`r round(mean(str_detect(study4_sample$authentic, fixed("personalised")), na.rm = TRUE) * 100)`%
of participants stated that the personalized AI was more authentic and 
`r round(mean(str_detect(study4_sample$effort, fixed("Carefully"))) * 100)`% of 
participants stated that giving the AI several detailed prompts was more 
effortful. This suggests that even if participants might not have felt the 
output was meaningfully authentic in the way that mattered (see Discussion), our 
participants agreed that using a personalized AI was at least more authentic 
than using a generic one.

### Pre-registration

We pre-registered the study on the Open Science Framework
(<https://osf.io/ac9g3/?view_only=912d9b57023d49baa87eea999574f0ce>).

### Statistical Analysis

We fitted the same Bayesian multivariate multilevel cumulative-link ordinal 
models as in Studies 1 and 2. All models converged normally ($\hat{R}$ ≤ 1.01).

## Results

We first looked across all the tasks. On average, we found that people who 
outsourced to AI in a low effort way were perceived as less competent, less 
moral, lazier, and less trustworthy than people who put more effort into their 
use of AI (@fig-treatments-study4; @tbl-treatment-diffs-study4). By contrast, we 
found that character evaluations did not differ between people who used a 
standard AI model rather than a personalized AI model. We also found no 
interaction effects between effort and the type of AI used.

```{r, fig.height=6, fig.width=7}
#| label: fig-treatments-study4
#| fig-cap: Overall Character Evaluations in Study 4
#| fig-align: center
#| apa-note: Participants in the control condition and four AI outsourcing conditions evaluated people in the scenarios on (a) competence, (b) warmth, (c) morality, (d) laziness, and (e) trustworthiness. Colored points represent participant responses to the questions, jittered for easier viewing. Black points are estimated marginal means from the fitted model, pooling over participants and tasks. Black points and line ranges represent posterior medians and 95% credible intervals, respectively.

tar_read(plot_treatments_study4)
```

::: {.landscape}

```{r}
#| label: tbl-treatment-diffs-study4
#| tbl-cap: Overall Pairwise Contrasts in Study 4 \vspace{20pt}
#| apa-note: Numbers reflect differences in marginal means on a 7-point Likert scale, pooling over participants and tasks. The bottom row represents the interaction between AI type and effort (i.e., the difference between the differences in the rows above). Main numbers are posterior medians, numbers in the square brackets are 95% credible intervals.

tar_read(table_treatment_diffs_effects_study4) %>%
  mutate(
    group = c(
      rep("Comparison to control", 4),
      rep("Effect of AI type", 2),
      rep("Effect of effort", 2),
      "Interaction effect"
    )
  ) %>%
  as_grouped_data(groups = "group") %>%
  as_flextable(hide_grouplabel = TRUE) %>%
  theme_apa() %>%
  width(width = c(3, rep(1.3, 5))) %>%
  bold(i = ~ !is.na(group)) %>%
  add_header_row(values = c("", "Response"), colwidths = c(1, 5)) %>%
  align(i = 1, j = NULL, align = "center", part = "header") %>%
  align(i = NULL, j = 1, align = "left", part = "all") %>%
  line_spacing() %>%
  fontsize(size = 10)
```

:::

As in Studies 1 and 2, the effects of outsourcing to AI varied across the 
different tasks, especially for perceptions of warmth and morality
(@fig-treatments-tasks-study4). We again found that the negative causal effects 
of outsourcing to AI were particularly strong for tasks that are social, require 
social skills, impact others, have important consequences, and require effort 
(Supplementary Figures [-@suppfig-interactions-study4] and 
[-@suppfig-interaction-pars-study4]). Indeed, for tasks like writing wedding 
vows or writing a love letter, outsourcing to a personalized AI in a high effort 
way was still perceived more negatively than the control condition for all five 
character dimensions.

::: {.landscape}

```{r, fig.height=5, fig.width=8.5}
#| label: fig-treatments-tasks-study4
#| fig-cap: Variation in the Effects of Outsourcing across Tasks in Study 4 \vspace{-30pt}
#| fig-align: center
#| apa-note: Tasks are ordered from most social (top) to least social (bottom) according to ratings from a pilot study. Point ranges are differences in marginal means on a 7-point Likert scale for the low effort conditions (red) and high effort conditions (blue) compared to the control condition. Upper panels refer to the standard LLM conditions, and lower panels refer to the personalized LLM conditions. Points and ranges represent posterior medians and 95% credible intervals, respectively.

tar_read(plot_treatments_tasks_study4)
```

:::

## Discussion

In Study 4, we found that effort is an important mechanism by which outsourcing 
to AI leads to negative character evaluations. People who engaged in effortless 
copying of the AI's first output were perceived more negatively than people who 
spent time and effort crafting the AI's output with multiple prompts. 
Nevertheless, for social tasks like writing wedding vows or love letters, 
outsourcing to AI in a high effort way was still perceived more negatively than 
completing the task oneself.

Interestingly, we found no effect of authenticity as proxied by the use of a 
personalized AI that is trained on one's own prior writings compared to a 
standard AI like ChatGPT. There are different explanations of this: it is 
possible that authenticity is not an important mechanism underlying the effect 
of outsourcing on negative character evaluations; it is possible that 
authenticity is important but our specific manipulation did not move the needle 
on authenticity enough to impact character evaluations; and it is possible that 
outsourcing to a personalized AI simply is not enough to overcome negative 
perceptions. While previous work has found an effect of personalized AI models 
on perceived credit [@Earp2024], and the majority of participants in our study 
stated that the personalized AI was more authentic than a standard model like 
ChatGPT, perceptions of *meaningful* authenticity in our study may have remained
low even with the personalized AI model. An AI could be perfectly trained on all
apologies that a person has ever written, but one might still think that a 
specific apology it then generates in a new instance is not an *authentic* 
apology. Therefore, even if people were described as outsourcing to an AI that 
was trained on their own writing and therefore personalized, participants still
may not have seen the specific output as being meaningfully authentic in the way
that matters for character judgments.

In Study 5, we turn to look at a third potential mechanism: a perceived lack of 
importance attached to the task. When participants read about someone who 
outsources to AI in our studies, they may be inferring that they simply did not 
care enough about the task -- "If this was important to them, they would do it 
themselves!". To the extent that we especially want people to care about their 
relationships with others -- the kind of things demonstrated through love 
letters, apology notes, and gift-giving -- this could explain the particular 
negativity we see for social tasks compared to tasks like writing daily 
schedules, recipes, or computer code. To test this, in Study 5, we attempted to 
counteract inferences about care for the task by explicitly telling participants 
that someone had a good reason for using AI: that they really cared about the 
task and used AI because they wanted to get it right.

# Study 5

## Methods

### Participants

```{r}
study5_sample <-
  tar_read(study5_data) %>%
  group_by(id) %>%
  summarise(
    gender = unique(gender),
    age = unique(age),
    chatgpt_used = unique(chatgpt_used)
    )
```

We used the same power estimate from Study 2 to determine our target sample size
of *n* = 750 (150 participants in each of five conditions). We recruited a 
convenience sample of 800 participants from the United Kingdom through Prolific. 
After excluding participants who failed our pre-treatment attention check, we 
were left with a final sample of `r nrow(study5_sample)` 
participants (`r sum(study5_sample$gender == "Female")` female; 
`r sum(study5_sample$gender == "Male")` male;
`r sum(study5_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(study5_sample$gender, "Prefer") | is.na(study5_sample$gender))`
undisclosed gender; mean age = `r round(mean(study5_sample$age), 2)` years).
`r round(mean(study5_sample$chatgpt_used == "Yes", na.rm = TRUE) * 100, 0)`% of
these participants reported having used ChatGPT before.

### Design

We used the same "control plus 2x2" between-subjects design as in Studies 1 and 
3. In the experimental conditions, we manipulated whether people in the 
scenarios used AI as a tool or "fully" outsourced to AI, and whether people had
bad or good reasons for using AI. This resulted in five conditions overall:
(*i*) the control condition, (*ii*) the tool-bad-reason condition, (*iii*) the 
tool-good-reason condition, (*iv*) the full-bad-reason condition, and (*v*) the 
full-good-reason condition.

### Procedure

The procedure was mostly identical to Study 4, with two changes. First, we 
reduced the number of tasks, focusing on eight tasks (four "social" tasks and 
four "non-social" tasks) that fit with the manipulation of the updated design 
(since, for example, participants might find it difficult to see how someone 
could deeply value a shopping list and want to get it right). Second, we
updated the presentation of the scenarios. We told participants in the
experimental conditions:

- *Bad reason conditions*: "Because they are really short on time and want to 
complete the task quickly, [the person] uses the AI tool ChatGPT."
- *Good reason conditions*: "Because this task is really important to them and 
they want to make sure they get it right, [the person] uses the AI tool
ChatGPT."

<br>

We then told participants:

- *Tool outsourcing conditions*: "[The person] asks ChatGPT to provide ideas, 
inspiration, and feedback, but they edit and rewrite the suggestions and finish
the task themselves."
- *Full outsourcing conditions*: "[The person] copies ChatGPT's output
word-for-word, rather than doing it themselves."

<br>

In addition to the five character evaluations, on each page we also asked 
participants, on a 7-point Likert scale, how much they thought the person cared
about the task.

### Pre-registration

We pre-registered the study on the Open Science Framework
(<https://osf.io/ac9g3/?view_only=912d9b57023d49baa87eea999574f0ce>).

### Statistical Analysis

We fitted the same Bayesian multivariate multilevel cumulative-link ordinal 
models as in Studies 1, 2, and 4. All models converged normally ($\hat{R}$ ≤ 
1.01).

## Results

We first looked across all the tasks. In line with our previous results, we 
found that people who fully outsourced to AI by copying the output verbatim were 
perceived as less competent, less moral, and less trustworthy than people who 
used AI as a collaborative tool (@fig-treatments-study5; 
@tbl-treatment-diffs-study5). Perhaps surprisingly, though, people's reasons for 
outsourcing to AI did not appear to influence character evaluations when pooling 
across all the tasks. When looking at the tasks overall, character evaluations 
did not differ between people who really cared about the task and wanted to get 
it right and people who used AI because they were short on time and wanted to 
complete the task quickly. This was true both when using the AI as a tool or 
outsourcing in full.

```{r, fig.height=6, fig.width=7}
#| label: fig-treatments-study5
#| fig-cap: Overall Character Evaluations in Study 5
#| fig-align: center
#| apa-note: Participants in the control condition and four AI outsourcing conditions evaluated people in the scenarios on (a) competence, (b) warmth, (c) morality, (d) laziness, and (e) trustworthiness. Colored points represent participant responses to the questions, jittered for easier viewing. Black points are estimated marginal means from the fitted model, pooling over participants and tasks. Black points and line ranges represent posterior medians and 95% credible intervals, respectively.

tar_read(plot_treatments_study5)
```

::: {.landscape}

```{r}
#| label: tbl-treatment-diffs-study5
#| tbl-cap: Overall Pairwise Contrasts in Study 5 \vspace{20pt}
#| apa-note: Numbers reflect differences in marginal means on a 7-point Likert scale, pooling over participants and tasks. The bottom row represents the interaction between outsourcing type and the reasons for outsourcing (i.e., the difference between the differences in the rows above). Main numbers are posterior medians, numbers in the square brackets are 95% credible intervals.

tar_read(table_treatment_diffs_effects_study5) %>%
  mutate(
    group = c(
      rep("Comparison to control", 4),
      rep("Effect of outsourcing type", 2),
      rep("Effect of reasons for outsourcing", 2),
      "Interaction effect"
    )
  ) %>%
  as_grouped_data(groups = "group") %>%
  as_flextable(hide_grouplabel = TRUE) %>%
  theme_apa() %>%
  width(width = c(2.5, rep(1.3, 6))) %>%
  bold(i = ~ !is.na(group)) %>%
  add_header_row(values = c("", "Response"), colwidths = c(1, 6)) %>%
  align(i = 1, j = NULL, align = "center", part = "header") %>%
  align(i = NULL, j = 1, align = "left", part = "all") %>%
  line_spacing() %>%
  fontsize(size = 10)
```

:::

Importantly, though, as in our previous studies, the type of task mattered 
(@fig-treatments-tasks-study5). Perceptions of outsourcing were particularly 
negative for tasks that are social, require social skills, impact others, have 
important consequences, and require effort (Supplementary Figures 
[-@suppfig-interactions-study5] and [-@suppfig-interaction-pars-study5]).
Indeed, for socio-relational tasks like writing an apology letter and writing 
wedding vows, people using AI as a tool for good reasons were still perceived 
more negatively than the control condition on the dimensions of warmth, 
morality, laziness, and care, though not on the dimensions of competence or 
trustworthiness.

::: {.landscape}

```{r, fig.height=4, fig.width=8.5}
#| label: fig-treatments-tasks-study5
#| fig-cap: Variation in the Effects of Outsourcing across Tasks in Study 5 \vspace{-30pt}
#| fig-align: center
#| apa-note: Tasks are ordered from most social (top) to least social (bottom) according to ratings from a pilot study. Point ranges are differences in marginal means on a 7-point Likert scale for the bad reason conditions (red) and good reason conditions (blue) compared to the control condition. Upper panels refer to the tool outsourcing conditions, and lower panels refer to the full outsourcing conditions. Points and ranges represent posterior medians and 95% credible intervals, respectively.

tar_read(plot_treatments_tasks_study5)
```

:::

Moreover, when we delved further into the task-specific estimates, we found that 
the reasons manipulation did indeed have an effect on character evaluations for 
social tasks -- but not non-social tasks (@suppfig-reasons-tasks-study5). When 
writing a bereavement card, for example, people were perceived as less warm, 
less moral, lazier, and less trustworthy when they used AI to save time compared 
to when they used it because they cared about doing the task well. The same was 
not true for non-social tasks like writing computer code or solving a 
mathematical equation.

## Discussion

In Study 5, we attempted to counteract the potential perception that outsourcing 
to AI reflects caring less about the task by explicitly informing participants 
about the person's reason for outsourcing: they outsourced to AI because they 
really cared about the task and wanted to get it right. As well as replicating 
our finding that fully outsourcing to AI is perceived more negatively than using 
AI as a tool, we also found an important effect of the reasons for outsourcing, 
but only for socio-relational tasks. When writing a bereavement card or an 
apology letter, for example, people were perceived more negatively if they used 
an AI tool to produce a quick output in a rush, rather than to ensure they got
it right. Nonetheless, for socio-relational tasks, the "best" use of AI in this 
study -- using AI as a tool because they cared about the task and wanted to get 
it right –- *still* led to targets being perceived more negatively than if they 
had completed the task themselves.

While we have so far shown varying evidence for three different mechanisms that 
might underlie the negative perceptions of outsourcing to AI -- effort, 
authenticity, and caring about the task -- it is likely that these mechanisms 
are related. For example, outsourcing to AI might indicate a lack of effort, 
which then might signal a lack of authenticity and reduced care in the task, 
leading to negative character evaluations. Our previous studies have been unable 
to test causal models like these as we manipulated the mechanisms separately and 
independently. In Study 6, therefore, we bring all three mechanisms together and 
test their combined associations with character evaluations. To do this, we 
focus on a single socio-relational task – writing a love letter – which we 
elaborate for participants with a more detailed vignette.

# Study 6

## Methods

### Participants

We conducted a power simulation to determine our target sample size. The 
simulation suggested that a sample size of 200 participants per condition 
(overall *n* = 600 for three conditions) would be required to detect a 
small-to-medium difference between conditions (Cohen's *d* ≈ 0.30) with above 
80% power.

```{r}
study6_sample <-
  tar_read(study6_data) %>%
  transmute(
    gender = gender,
    age = age,
    chatgpt_used = chatgpt_used,
    comp1 = comprehension1 == "A love letter",
    comp2 = ifelse(treatment == "Control", comprehension2 == "No", 
                   comprehension2 == "Yes")
    )
```

We recruited a convenience sample of 651 participants from the United Kingdom 
through Prolific. After excluding participants who failed our pre-treatment 
attention check, we were left with a final sample of `r nrow(study6_sample)` 
participants (`r sum(study6_sample$gender == "Female")` female; 
`r sum(study6_sample$gender == "Male")` male;
`r sum(study6_sample$gender == "Non-binary / third gender")`
non-binary / third gender;
`r sum(str_starts(study6_sample$gender, "Prefer") | is.na(study6_sample$gender))`
undisclosed gender; mean age = `r round(mean(study6_sample$age), 2)` years).
`r round(mean(study6_sample$chatgpt_used == "Yes", na.rm = TRUE) * 100, 0)`% of
these participants reported having used AI tools like ChatGPT before.

### Design

We randomly allocated participants into one of three conditions in a 
between-subjects design: (*i*) the control condition, (*ii*) the tool 
outsourcing condition, or (*iii*) the full outsourcing condition. These 
conditions determined how the scenario was presented to participants.

### Procedure

We presented participants with a vignette about a person, Adam, who is writing a
love letter in a Valentine's Day card to his partner (see Supplementary 
Materials for full vignette wording). We told participants in each of the 
conditions:

- *Control condition*: "Adam decides to write the love letter in the card by 
himself."
- *Tool outsourcing condition*: "Adam decides to use AI to help write the love
letter in the card. He asks ChatGPT to provide ideas, inspiration, and feedback,
but he edits and rewrites the suggestions and finishes writing the love letter
himself."
- *Full outsourcing condition*: "Adam decides to use AI to write the love letter
in the card. He asks ChatGPT to write the love letter and copies the output
word-for-word, rather than writing it himself."

<br>

We then presented participants with the love letter that Adam wrote (in reality,
this was written by ChatGPT version 4o; see Supplementary Materials for
wording). On the following page, we asked participants what Adam wrote and 
whether he used AI to help. 
`r round(mean(study6_sample$comp1 & study6_sample$comp2, na.rm = TRUE) * 100)`%
of participants answered both of these comprehension questions correctly.

Using 7-point Likert scales, we then asked participants how much effort they 
thought Adam put into the love letter, how authentic they thought the love 
letter was, how much they thought Adam cared about the love letter, and the same
five character evaluations as in our previous studies. In additional free
response questions, we asked participants to explain how they felt towards Adam
and how they would feel if they were Adam's partner. Finally, we asked
participants several questions about AI tools like ChatGPT.

### Pre-registration

We pre-registered the study on the Open Science Framework
(<https://osf.io/ac9g3/?view_only=912d9b57023d49baa87eea999574f0ce>).

### Statistical Analysis

We fitted two Bayesian regression models to the data. The first model was a 
multivariate cumulative-link ordinal model including all Likert scales as 
separate response variables. The second model was a path model capturing the 
effect of outsourcing on character evaluations, both directly and indirectly 
through perceptions of effort, authenticity, and care. In this second model, we
included ordinal predictors as monotonic effects and modelled the five character
evaluations as a single latent variable. We used regularizing priors for all
parameters to impose conservatism on parameter estimates. All models converged 
normally ($\hat{R}$ ≤ 1.01).

## Results

Across all measures, we found that outsourcing the love letter to AI was 
perceived more negatively compared to the control condition and that fully 
outsourcing to AI was perceived more negatively than using AI as a collaborative
tool (@fig-treatments-study6; @tbl-treatment-diffs-study6). Not only did 
outsourcing the love letter lead to more negative character evaluations, but 
outsourcing to AI was also seen as less effortful, less authentic, and 
indicative of caring less about the task.

```{r, fig.height=5, fig.width=7}
#| label: fig-treatments-study6
#| fig-cap: Perceptions of the Person and the Love Letter in Study 6
#| fig-align: center
#| apa-note: Participants in the control, tool outsourcing, and full outsourcing conditions rated (a) the amount of effort put into the love letter, (b) how authentic the love letter was, (c) how much the person cared about the love letter, and (d-h) five character evaluation measures. Colored points represent participant responses to the questions, jittered for easier viewing. Black points are estimated marginal means from the fitted model. Black points and line ranges represent posterior medians and 95% credible intervals, respectively.

tar_read(plot_treatments_study6)
```

::: {.landscape}

```{r}
#| label: tbl-treatment-diffs-study6
#| tbl-cap: Pairwise Contrasts in Study 6 \vspace{20pt}
#| apa-note: Numbers reflect differences in marginal means on a 7-point Likert scale. Main numbers are posterior medians, numbers in the square brackets are 95% credible intervals.

tar_read(table_treatment_diffs_study6) %>%
  flextable() %>%
  theme_apa() %>%
  width(width = c(rep(1.1, 9))) %>%
  add_header_row(values = c("", "Response"), colwidths = c(1, 8)) %>%
  align(i = 1, j = NULL, align = "center", part = "header") %>%
  align(i = NULL, j = 1, align = "left", part = "all") %>%
  line_spacing() %>%
  fontsize(size = 8)
```

:::

Exploratory text analysis of participants' free responses supported this 
quantitative pattern (see Supplementary Materials for methodology and 
@supptbl-text-analysis-study6 for results). When comparing word frequencies 
between conditions, we found that Adam was more likely to be described as "lazy"
and less likely to be described as "caring", "thoughtful", and "genuine" in both 
outsourcing conditions compared to the control condition. Adam was also more 
likely to be described as "romantic" and as someone who "loves" his partner when
he used AI as a collaborative tool, compared to when he fully outsourced the
love letter to AI.

When we included all the variables in a single path model, we found that 
outsourcing influenced character evaluations both directly and indirectly 
through our proposed mechanisms (@fig-path-model-study6). The indirect effects 
showed that people perceived outsourced work as less effortful, and less 
effortful work was seen as less authentic and indicating less care about the 
task. In turn, less authenticity and care were associated with more negative 
evaluations of the person. Effort itself was not directly related to character 
evaluations, suggesting that effort works solely through perceptions of 
authenticity and care.

```{r, fig.height=4.5, fig.width=7.5}
#| label: fig-path-model-study6
#| fig-cap: Path Model in Study 6
#| fig-align: center
#| apa-note: All predictors were modelled as monotonic effects, such that parameters can be interpreted as the expected average difference between two adjacent categories of the ordinal predictor on the logit scale. The "evaluations" outcome variable was modelled as a single latent variable with loadings from all five character evaluations (competence, warmth, morality, laziness, and trustworthiness).

tar_read(plot_path_model_study6)
```

<br>

# General Discussion

In a world where openly available generative AI LLMs promise to let people do 
more tasks, more efficiently, what is the value of doing something oneself? 
People can – and *do* – use AI tools like ChatGPT to, for example, create 
dinner recipes, assist with coding, and even write job applications 
[@UKGov2024]. But it is not only such routine, everyday, and non-social tasks 
that AI now "assists" with. People can use AI for a seemingly endless range of 
more social and consequential tasks too, from crafting apology letters to 
writing condolences to even writing wedding vows. But does it matter whether 
people do things themselves or instead "efficiently" reduce their effort -- and
if so, why? In this paper, across six pre-registered experiments, we show how 
and why AI-outsourcing shapes perceptions of others in a world where outsourcing 
has never been easier and cheaper.

Psychologists have long been interested in the apparent paradox of effort: it is 
both costly and valued [see @Inzlicht2018]. In more recent years, scholars have 
turned to consider not only why people might value effort itself, but how (not) 
expending effort shapes how we are viewed by others. @Celniker2023 argue that 
effort is moralized because people seem to infer positive moral character and 
cooperative intent from effort. But yet it has been unclear how outsourcing 
itself, and not reduced effort per se, may shape judgments, especially when
people outsource to artificial agents. Moreover, previous work looking at effort 
has tended to focus on a small number of examples that are not the most socially 
impactful: looking at how people judge the quality of products like paintings, 
poems, or suits of armor based on how much effort is perceived [@Kruger2004] or 
looking at how people judge others who expend more or less effort in employment 
and physical exercise [@Celniker2023] or, in the context of AI, those who use AI 
for work [@Reif2025]. But it has remained unclear how we might view others who 
outsource to AI; how these effects might vary based on how socio-relational the 
task is; how different ways of outsourcing and expending less effort influence 
perceptions; and how outsourcing has different effects on different kinds of 
social perceptions. And perhaps most importantly, through asking these 
questions, we gain insight into the most important question: why exactly does 
effort have the effects that it does? Is it that expending less effort directly 
leads to negative social perceptions, as some have argued in psychology 
[@Celniker2023] or, more in line with philosophical perspectives 
[e.g., @Aristotle2009] might expending effort also lead to second-order 
judgments about how much someone values a task, which then independently drives 
negative perceptions? Across our studies, we shed new light on all of these 
questions, providing a greater understanding of how, when, and why doing
something oneself the hard way shapes character judgments.

In Study 1, we showed that people who expended less effort by outsourcing tasks 
to AI or other humans were perceived more negatively than people who completed 
the tasks by themselves. These negative impressions were seen across ratings of 
competence, warmth, morality, laziness, and trustworthiness, and the effects 
were particularly strong for people who used AI to complete socio-relational 
tasks, such as writing a love letter or writing wedding vows. In Study 2, we 
focused in on AI outsourcing specifically and replicated the results of Study 1,
showing again that negative impressions were particularly strong for people who 
used AI to complete socio-relational tasks, and were found even for the "best 
case" of openly acknowledging the use of AI as a collaborative tool. In Study 3, 
we showed that people perceive both the outsourcer and the outsourced work more 
negatively, with outsourced work perceived as less meaningful, less authentic, 
and less reward-worthy than ostensibly human-generated writing. In Study 4, we 
showed that while it matters whether people spent time crafting the AI prompts 
or simply gave a rushed initial prompt, even expending effort into crafting the 
best prompts was still not enough to counteract the negative effects from using 
AI. In Study 5, we explored the potential role of the reasons for using AI and 
found that while explicitly telling participants that the person used AI because 
they cared about the task reduced negative perceptions for social tasks, it was 
still not enough to eliminate negative perceptions completely. In Study 6, we 
showed that a perceived lack of effort is taken to signal both a lack of 
authenticity and lack of importance attached to the task, and these 
independently influenced character judgments above and beyond the effect of 
effort.

Together, these findings advance our understanding of how effort shapes 
character judgments in an AI-mediated world. We show that people negatively 
judge those who outsource to AI, just as we negatively judge those who outsource 
to humans. We show that the type of task does matter: outsourcing 
socio-relational tasks to AI (and humans) leads to particularly negative 
perceptions, especially for perceptions of warmth and morality. We show that 
different ways of AI outsourcing lead to differences in the degree of negative 
perceptions but that, critically, even outsourcing to AI in the "best" way 
(e.g., using it as a tool and finishing the work oneself while being honest 
about the AI use) is still not enough to eliminate the negative consequences. 
And finally, advancing theoretical work on effort, we provide further insight 
into *why* effort matters. In contrast to theoretical models in which effort 
itself directly signals positive moral character, our work here suggests that 
reduced effort also leads to second order judgments that independently drive 
negative judgments. The reduced effort from outsourcing socio-relational tasks 
to AI signals that the work is less authentically one's own and that the person 
cares less about the task (and therefore, perhaps, the relationship). We show 
that people infer that when someone outsources they care less about the task, 
and even directly telling participants the person cared was not enough to 
eliminate this perception for social tasks. Moreover, the lack of a direct 
effect of perceived effort in our path model showed that it is inferences of 
authenticity and care, rather than perceived effort per se, that are associated 
with negative character evaluations. As a participant in our final study put it: 
"*If he really cared, he would have just done it by himself from scratch*" 
(female, 25 years old).

Our findings cohere with the philosophical idea that there is value in *how* a 
task was done, and not merely *whether* it was done [@Aristotle2009; 
@Goodman2010; @Hursthouse2023; @Stohr2006]. For many socio-relational tasks, it 
might seem that part of the constitutive action is the *process* by which it 
occurs: an apology that does not contain a genuine reflection and commitment to 
do better, rather than just the words "I am sorry", might not seem to be an 
apology at all. In contrast, for many of the non-social tasks, it is easier to 
distinguish the importance of the process from the outcome. In this way, our 
work suggests that people rarely adopt a purely utilitarian perspective in which 
outcomes are the sole determinant [@Everett2020; @Kahane2018]. Instead, their 
judgments cohere more with ideas from virtue ethics about the importance of 
*doing* [@Hursthouse2023; @Stohr2006]. Outsourcing to AI -- especially for 
social tasks –- may allow us to produce similar outputs, but by severing the 
outcome from the practice of doing, it may risk the development and maintenance
of our human virtues [@Vallor2015; @Vallor2024].

AI is often being marketed as being able to help us to do more and more tasks, 
promising gains of efficiency that align with societal incentives for "hacks" 
that encourage people to do more with less energy and effort. Our work, 
however, highlights that when it comes to our psychology, efficiency is not the 
only currency. Instead, *in*efficiency can sometimes pay off more, especially 
for social tasks. By expending effort themselves instead of outsourcing to AI, 
people are able to signal authenticity and care for the task, and this can lead 
to better reputations [see also @Celniker2023]. Correspondingly, expending 
effort, even "unnecessarily", is not as irrational, biased, or suboptimal as we 
might think from a utilitarian perspective in which outcomes are the only things 
that matter. Instead, it is precisely this inefficiency that helps people signal 
things that they care about and connect with others, thereby arguably reflecting 
a deeply rational reflection of virtues and the importance of social ties
[@Everett2016].

Most speculatively, our results on the negative effects of AI-outsourcing on 
character judgments highlight potential risks in how increased use of AI could 
lead to negative consequences for social ties, especially if people start to 
assume, by default, that others are using AI for the kind of tasks that matter. 
Sociologists have highlighted concerns about the negative effects that 
outsourcing to AI can have on our "connective labor", arguing that while AI can 
enhance certain tasks, it cannot replicate the depth of human relationships 
essential for effective caregiving, education, and support [@Pugh2024]. Similar 
arguments have been made about the risks of outsourcing empathy to AI 
[@Landes2025]. In this way, the rapid move towards using AI for more and more 
tasks could have serious and unintended consequences on the way we connect with 
one another, serving to further weaken the social ties that bind us into a 
community.

## Limitations and Directions for Future Research

The studies in this paper are not without their limitations. While we included a 
range of different socio-relational and professional tasks in an effort to 
improve the generalizability of our findings across domains, it would be 
interesting for future work to additionally explore the generalizability and 
variability of our findings across countries with different AI infrastructures 
and readiness levels [@OxfordInsights; @TortoiseMedia] and over time as AI use 
becomes more commonplace. By focusing on generalizability across various 
real-world tasks in which people outsource, it could also be argued that our 
design lacks the richness of information in extended vignettes that might 
influence character evaluations. Building on work understanding effort as being
intrinsically moralized, we have advanced previous psychological theory by 
highlighting the *ways* in which effort influences perceptions of authenticity 
and care. It will be interesting for future research to delve deeper into these 
proposed mechanisms, both philosophically and psychologically: *why* is it that 
the perceived care for the task matters, and what are the boundary conditions of 
these effects? Finally, while we have demonstrated negative perceptions of 
outsourcing in this paper, it will be important for future research to explore 
when people might deem outsourcing to AI as acceptable or even preferable. 
Several of the participants in our final study expressed in their free responses 
that they would have been okay with Adam using AI to write the love letter if he 
was not a confident writer or had a learning difficulty that made writing 
challenging, such as dyslexia. In line with this, some research has found that 
people are more accepting of cognition-enhancing technologies and drugs when 
they are used to repair cognitive functions, rather than to enhance cognitive 
functions beyond "normal" levels [@Medaglia2019; @Rudski2014]. Future research 
should explore whether negative perceptions of outsourcing persist when AI is 
used in a reparative way.

## Conclusions

To conclude, across six pre-registered studies, we have demonstrated negative 
perceptions of outsourcing to AI. Our participants perceived individuals who 
outsource tasks to AI more negatively across a range of character dimensions and 
perceived outsourced work as less meaningful and authentic. Negative perceptions 
were particularly strong for socio-relational tasks, such as writing wedding 
vows, and were compounded when the outsourcer copied the AI’s output verbatim 
and did not honestly acknowledge their use of AI. Our findings advance previous 
theoretical work, connect with broader debates about the importance of *doing* 
in social relationships, and highlight that for many tasks -- especially those 
that are more socio-relational -- it might be better to move away from a focus 
on making things more efficient at all costs and instead bring back a 
recognition of the power of *in*efficiency. Doing something oneself, even if AI 
could do it quicker and easier, signals one that is authentic and cares about 
the task and therefore can help bind us together. In a world of 
algorithm-mediated interactions, AI is no substitute for investing effort into 
our interpersonal relationships.

\newpage

# Acknowledgements

This work was generously supported by funding from REDACTED.

# Data and Code Availability

All data and original code can be found here:
<https://osf.io/ac9g3/?view_only=912d9b57023d49baa87eea999574f0ce>.

# Statement of Interests

The authors have no conflicts of interest to disclose.

\newpage

# References

<!-- References will auto-populate in the refs div below -->

::: {#refs}
:::

\newpage

{{< include appendix.qmd >}}